
% APF v51 -- Bottom/Top/Charm Shared QCD Covariance Ledger
\section{Bottom--Top--Charm Shared QCD Covariance Ledger}
\label{sec:heavy-qcd-ledger-v51}

\paragraph{Purpose.}
Versions v37, v50, and the inherited bottom line now give three heavy-quark
export-candidate routes.  This section closes the shared QCD ledger discipline
that prevents those routes from being compared with mixed codomains or tuned by
post-hoc choices of $\alpha_s$, thresholds, loop order, or scheme conversion.

\begin{theorem}[Shared heavy-quark QCD ledger admissibility]
Let the heavy-quark route registry be
\[
 c:\ m_c^{\overline{\rm MS}}(m_c),\qquad
 b:\ m_b^{\overline{\rm MS}}(m_b),\qquad
 t:\ m_t^{\rm MSR}(R_\star).
\]
A joint heavy-quark export claim is APF-admissible only if it carries the typed
QCD ledger
\[
\mathcal L_{\rm QCD}^H=
\{S_H,\mu_H,\alpha_s, L_\beta,L_{\gamma_m},\mathcal T,\mathcal M,
\mathcal C,\Sigma_{\rm QCD}\},
\]
where $S_H$ is the scheme family, $\mu_H$ the route scale, $\mathcal T$ the
threshold/matching convention, $\mathcal M$ the scheme-conversion map,
$\mathcal C$ the constants ledger, and $\Sigma_{\rm QCD}$ the propagated joint
covariance.  None of these ledger entries may be chosen from the residuals of
$c,b,t$ after comparison.
\end{theorem}

\paragraph{Codomain-aligned audits.}
The self-scale $\overline{\rm MS}$ branches admit same-codomain external audits:
\[
\Phi_c^{v50}=1.272334177712\ {\rm GeV},\qquad
m_c(m_c)_{\rm PDG2025}=1.2730\pm0.0046\ {\rm GeV},
\]
so
\[
\Delta_c=-0.000665822288\ {\rm GeV},
\qquad \Delta_c/\sigma_c=-0.144743975597.
\]
For bottom,
\[
\Phi_b^{v51}=m_b^{\rm TRACE}=4.177490455927\ {\rm GeV},\qquad
m_b(m_b)_{\rm PDG2025}=4.183\pm0.007\ {\rm GeV},
\]
so
\[
\Delta_b=-0.005509544073\ {\rm GeV},
\qquad \Delta_b/\sigma_b=-0.787077724714.
\]
Both same-codomain audits lie inside the quoted external comparison
uncertainties.

\paragraph{Top boundary.}
The top branch is not a same-codomain comparison with the direct, pole, or MC
masses.  Its closed APF object remains
\[
 m_t^{\rm MSR}(R_\star),\qquad R_\star=85.8572226983853\ {\rm GeV}.
\]
A comparison to pole, MC, or $\overline{\rm MS}$ top masses requires an explicit
MSR-conversion ledger and covariance.  Without that ledger, a numerical top
comparison is blocked rather than approximated.

\paragraph{Status.}
Version v51 promotes the heavy-quark sector to
\[
\boxed{
(c,b):[P_{\rm export\ candidate}^{\overline{\rm MS}\ {\rm self-scale}+QCD\ ledger}],
\qquad
 t:[P_{\rm export\ candidate}^{{\rm MSR}(R_\star)+QCD\ ledger}].
}
\]
It does not claim pole masses, MC masses, zero residuals, full physical-final
masses, or light-quark chiral/lattice closure.
