
\section{Charged-Lepton QED Closure Boundary Theorem (v49)}

\subsection{Purpose}
The preceding layers closed the charged-lepton APF pole route and typed the QED running transport.  This section records the final closure boundary.  The aim is not to tune the last diagnostic residuals to zero, but to decide exactly what has been completed and what cannot be completed without importing an external QED coefficient, threshold, matching, and covariance ledger.

\begin{theorem}[Charged-lepton QED closure boundary]
Let
\[
\Phi_\ell^{v43}=(0.511002635788531,\;105.658243985342,\;1776.916832008411)\ {\rm MeV}
\]
be the APF charged-lepton pole vector obtained from the APF trace vector, generation residual operator, APF count coefficients, scalar normalization, and APF QED covariance envelope.  Let the diagnostic external ledger be the CODATA 2022/NIST charged-lepton mass-energy ledger and inverse fine-structure constant
\[
\alpha^{-1}=137.035999177(21).
\]
Then the charged-lepton QED branch is closed in the following precise sense:
\[
\boxed{
\ell:\ [P_{\rm pole\ completion}^{\rm APF\ envelope,CODATA2022}]
+[P_{\rm LO\ QED\ numeric\ witness}]
+[P_{\rm QED\ truncation\ covariance\ admitted}]
+[P_{\rm full\ QED\ ledger\ boundary}].
}
\]
The remaining full-running numerical map is not an APF-side missing theorem; it is an external-ledger problem requiring a declared scheme, scale, loop order, matching convention, active-threshold schedule, coefficient list, and covariance matrix.
\end{theorem}

\subsection{Numerical closure audit}
The CODATA-ledger residuals of the APF pole vector are
\[
\delta_\ell=(+0.000721155792\%,\;-0.000124471588\%,\;+0.003198451674\%).
\]
The APF envelope is
\[
\varepsilon_{\rm APF}=100/5063=0.019751135690302\%,
\]
and the declared perturbative QED truncation scale is
\[
\varepsilon_{\rm QED}^{(2)}=100(\alpha/\pi)^2=0.000539549026324\%.
\]
Thus the combined non-fit envelope is
\[
\varepsilon_{v49}=\sqrt{\varepsilon_{\rm APF}^2+(\varepsilon_{\rm QED}^{(2)})^2}
=0.019758503845396\%.
\]
The maximum normalized residual is
\[
\max_i |\delta_i|/\varepsilon_{v49}=0.161877219998<1.
\]
Therefore all three charged leptons remain inside the declared APF+QED covariance envelope after official ledger import.

\subsection{One-loop transport witness}
The one-loop self-scale numerical witness remains
\[
\bar m_i^{\rm LO}(\mu_i=m_i)=\frac{\Phi_{\ell,i}^{v43}}{1+\alpha/\pi},
\]
with
\[
\frac{1}{1+\alpha/\pi}=0.997682563522362.
\]
Hence
\[
\bar m_\ell^{\rm LO}=(0.509818419640,\;105.413387716567,\;1772.798940124186)\ {\rm MeV}.
\]
This is a reproducible transport witness.  It is not promoted to a final multi-loop running-mass theorem.

\subsection{No-smuggling closure}
The following are not used as APF source inputs: CODATA masses for constructing the trace vector, inverse reconstruction of \(T_\ell\), target-selected values of \(\alpha\), target-selected renormalization scales, target-selected loop order, generation-specific QED correction factors, least-squares cleanup, or residual-dependent threshold schedules.  QED corrections are universal route objects once a scheme and ledger are declared; they are not three independent generation knobs.

\subsection{External-ledger boundary}
A final numerical QED-running theorem would require the external object
\[
\mathcal L_{\rm QED}^{\rm full}=\{S_{\rm run},\mu,L,K_1,\dots,K_L,\mathcal T,\mathcal M,\alpha(\mu),\Sigma_{\rm QED}\}.
\]
Without this object, the APF framework forbids promotion to
\([P_{\rm QED\ running\ final}]
for the same reason that it forbids hidden residual fitting: the target codomain would not be a fully declared continuation context.  With this object supplied, the task is downstream metrological transport, not APF charged-lepton source closure.

\subsection{Final claim ladder after v49}
\[
\boxed{
[P_{\rm QED\ branch\ complete\ at\ APF\ pole\ plus\ external\ ledger\ boundary}]
}
\]
means that the APF-side charged-lepton source, pole codomain completion, one-loop QED witness, perturbative covariance admission, and full-running ledger boundary are closed.
It does not mean
\[
[P_{\rm QED\ running\ final}],\qquad [P_{\rm physical\ final}],\qquad \Delta_\ell=0.
\]
