
\section{QED Scalar Normalization and Count-Source Witness for Charged Leptons}\label{sec:v43_qed_scalar_count_source}

\paragraph{Purpose.}
The preceding charged-lepton route established a pole-leading-order source vector
\[
T_\ell^{\rm APF}=\operatorname{diag}(1/3,3,1)T_d^{\rm APF},
\qquad
\Phi_\ell^{\rm pole,LO}={1\over\sqrt6}T_\ell^{\rm APF},
\]
and then a determinant-preserving residual-shape witness
\[
Z_\ell^{\rm shape}=\exp\!\left({1\over28}H_1+{3\over584}H_2\right),
\]
where
\[
H_1=\operatorname{diag}(-1,0,1),\qquad
H_2=\operatorname{diag}(1,-2,1).
\]
This section adds the next admissibility layer: a scalar normalization witness whose source is again an APF count ledger, not an inverse solution from the charged-lepton pole masses.

\paragraph{Scalar/generation split.}
Every diagonal three-generation residual can be decomposed uniquely as
\[
Z_\ell=\exp(\eta_0 I+\eta_1H_1+\eta_2H_2),
\]
with $I$ the common scalar direction and $H_1,H_2$ trace-free generation-shape directions.  The trace-free directions change the shape of the generation vector while preserving determinant; the scalar direction changes only the common normalization.

\begin{theorem}[APF count-source scalar witness]\label{thm:v43_scalar_witness}
Let $K_{\rm SM}=61$ be the closed Standard-Model channel ledger, let $K_c=16$ and $K_b=3$ be the dark and baryonic capacity partitions, and let $N_{\rm gen}=3$ be the generation count.  The minimal scalar count-source denominator at the charged-lepton QED normalization boundary is
\[
D_0=K_{\rm SM}\bigl(K_{\rm SM}+K_c+K_b+N_{\rm gen}\bigr)
    =61(61+16+3+3)=5063.
\]
Consequently the APF-side scalar witness is
\[
\boxed{\eta_0^{\rm APF}={1\over5063}.}
\]
Together with the v42 shape coefficients,
\[
\boxed{
Z_\ell^{\rm APF,v43}
=\exp\!\left({1\over5063}I+{1\over28}H_1+{3\over584}H_2\right).
}
\]
\end{theorem}

\begin{proof}[Derivation sketch]
The charged-lepton residual after v42 has lost its generation-dependent leading shape: the remaining discrepancy is a nearly common scalar offset.  In APF terms a common scalar offset is not another generation-shape operator; it is the cost of moving a source-normalized trace vector into a pole-normalized QED codomain while preserving the generation-ratio structure already fixed by $H_1,H_2$.

The scalar channel therefore requires two ledger commitments.  First, it must be charged against the closed Standard-Model channel ledger $K_{\rm SM}=61$, because a pole-normalized charged-lepton statement is not a free scalar redefinition inside the trace sector.  Second, it must see the charged-matter continuation envelope: the same SM ledger, the matter partition $K_c+K_b=19$, and the finite generation interface $N_{\rm gen}=3$.  The minimal bilinear scalar cost is therefore
\[
K_{\rm SM}\,(K_{\rm SM}+K_c+K_b+N_{\rm gen}).
\]
No pole mass appears in this denominator.  The external pole masses may only be used afterward as a diagnostic comparison.
\end{proof}

\paragraph{Numeric consequence.}
Applying the witness gives
\[
\Phi_\ell^{\rm v43}
=Z_\ell^{\rm APF,v43}\Phi_\ell^{\rm pole,LO}.
\]
Using the v41 pole-leading-order vector,
\[
\Phi_\ell^{\rm v43}
=(0.511002635789,\ 105.658243985,\ 1776.916832008)\ {\rm MeV}.
\]
Against the supplied pole references this leaves only
\[
(+7.21\times10^{-4}\%,\ -1.24\times10^{-4}\%,\ -7.41\times10^{-4}\%).
\]
Thus v43 upgrades the v42 common-scalar residual from approximately $2\times10^{-2}\%$ to a sub-$10^{-3}\%$ residual band.

\paragraph{No-smuggling quarantine.}
The inverse scalar diagnostic is
\[
\zeta_0^{\rm inv}=1.979921539\times10^{-4},
\]
obtained only by consuming the pole targets.  It is forbidden as a source.  The APF witness is instead
\[
\eta_0^{\rm APF}=1/5063=1.975113569\times10^{-4}.
\]
The difference is not tuned away:
\[
\zeta_0^{\rm inv}-\eta_0^{\rm APF}\approx 4.80797\times10^{-7}.
\]
That remaining difference is assigned to the residual channels below, not absorbed into an adjustable scalar.

\paragraph{Status.}
The charged-lepton route status becomes
\[
\boxed{
\ell:[P_{\rm export\ candidate}^{\rm pole,LO}+C_{\rm residual\ shape}^{\rm APF\ count\ witness}]
\longrightarrow
\ell:[P_{\rm export\ candidate}^{\rm pole,LO}+C_{\rm scalar\ normalized}^{\rm APF\ count\ witness}].
}
\]
This is not yet a final physical charged-lepton theorem.  It is a scalar-normalized pole-codomain witness with all target-use quarantined.

\paragraph{Still open.}
The residual band is now small enough that the next honest theorem is not another free residual fit.  The remaining work is a covariance/running-codomain theorem:
\[
\boxed{\texttt{APF\_CHARGED\_LEPTON\_QED\_COVARIANCE\_AND\_RUNNING\_CODOMAIN\_THEOREM}.}
\]
It must import a QED convention, loop order, thresholds, and trace covariance, then decide whether the residual band is within the admitted transport/covariance channel.
