\section{Charged-Lepton Full QED Running Ledger and Covariance Evaluation (v48)}
\label{sec:charged-lepton-full-qed-v48}

\paragraph{Status.}
This section records the v48 completion of the QED-running \emph{ledger} layer for the charged-lepton branch. It does not convert the branch into a final multi-loop running-mass prediction. The status stamp is
\[
\boxed{\texttt{CHARGED\_LEPTON\_FULL\_QED\_RUNNING\_LEDGER\_AND\_COVARIANCE\_EVALUATION\_PASS}}.
\]

\begin{definition}[QED running ledger]
A numerical QED running export is admissible only after the external ledger
\[
\mathcal L_{\rm QED}
=\{S_{\rm run},\mu,L,\alpha,\mathcal T,\mathcal M,\mathcal K_1,\ldots,\mathcal K_L,\Sigma_{\rm QED}\}
\]
is fixed before comparison with the target charged-lepton masses. Here $S_{\rm run}$ is the running-mass scheme, $\mu$ is the scale convention, $L$ is the loop order, $\mathcal T$ is the threshold schedule, $\mathcal M$ is the matching convention, $\mathcal K_n$ are the perturbative coefficients in that convention, and $\Sigma_{\rm QED}$ is the covariance protocol.
\end{definition}

\begin{theorem}[Full-running ledger typing]
The charged-lepton QED transport operator has the typed form
\[
U_{\rm QED}^{{\rm pole}\to S_{\rm run}}(\mathcal L_{\rm QED})
=\exp\!\left(\sum_{n=1}^L
\mathcal K_n(\mathcal L_{\rm QED})
\left({\alpha\over\pi}\right)^n\right),
\]
and the numerical export
\[
\Phi_\ell^{S_{\rm run}}(\mu)
=U_{\rm QED}^{{\rm pole}\to S_{\rm run}}(\mathcal L_{\rm QED})\,\Phi_\ell^{v43}
\]
is APF-admissible only if every entry of $\mathcal L_{\rm QED}$ is declared independently of the target masses.
\end{theorem}

\begin{proof}[APF proof sketch]
By the trace-to-scheme export theorem, export is an admissible continuation from a trace context to a measurement-scheme context. A running charged-lepton mass is not a scalar target; it is a scheme- and scale-indexed object. Therefore the codomain, route, constants, covariance, no-smuggling audit, and residual channels must all be fixed. The entries of $\mathcal L_{\rm QED}$ are precisely these commitments for the QED transport layer. If any of them are chosen after seeing the residual vector, the map ceases to be a source-independent continuation and becomes inverse fitting.
\end{proof}

\paragraph{Frozen CODATA 2022 ledger used for v48 diagnostics.}
The external constants used in this diagnostic layer are
\[
\alpha^{-1}=137.035999177(21),
\qquad
\alpha=0.007297352564331424,
\qquad
{\alpha\over\pi}=0.002322819464195329.
\]
The diagnostic mass-energy references are
\[
(m_e,m_\mu,m_\tau)_{\rm CODATA\ 2022}
=(0.51099895069,105.6583755,1776.86)\ {\rm MeV}.
\]
These are external references and transport constants only; they are not used to construct the APF source vector.

\paragraph{Inherited APF pole vector.}
The APF charged-lepton pole-codomain vector is
\[
\Phi_\ell^{v43}
=(0.511002635789,105.658243985342,1776.916832008411)\ {\rm MeV}.
\]
The relative residuals against the CODATA 2022 diagnostic ledger are
\[
\delta_\ell
=(+0.000721155792\%,-0.000124471588\%,+0.003198451674\%).
\]

\paragraph{QED perturbative scale.}
The evaluated one-loop self-scale witness remains
\[
\bar m_i^{\rm LO}(\mu_i=m_i)
={\Phi_{\ell,i}^{v43}\over 1+\alpha/\pi},
\qquad
{1\over 1+\alpha/\pi}=0.997682563522362.
\]
Thus
\[
\bar m_\ell^{\rm LO}
=(0.509818419640,105.413387716567,1772.798940124186)\ {\rm MeV}.
\]
The next perturbative scale, without importing any convention-dependent coefficient, is
\[
100\left({\alpha\over\pi}\right)^2
=0.000539549026324\%.
\]
This number is not used to tune any generation. It is a declared order-of-magnitude truncation scale attached to the QED route.

\begin{theorem}[v48 covariance admission]
Let
\[
\varepsilon_{\rm APF}={100\over5063}=0.019751135690302\%,
\qquad
\varepsilon_{\rm tr}=100\left({\alpha\over\pi}\right)^2=0.000539549026324\%.
\]
Define the conservative combined envelope
\[
\varepsilon_{\rm v48}
=\sqrt{\varepsilon_{\rm APF}^2+\varepsilon_{\rm tr}^2}
=0.019758503845396\%.
\]
Then each CODATA-ledger residual satisfies
\[
|\delta_i|<\varepsilon_{\rm v48},
\]
with maximum normalized residual
\[
\max_i {|\delta_i|\over\varepsilon_{\rm v48}}
=0.161877219998<1.
\]
Therefore the charged-lepton pole route is admitted by the APF+QED perturbative covariance envelope.
\end{theorem}

\begin{proof}
Direct substitution gives the residual vector displayed above. The envelope $100/5063$ is APF-derived from the v43--v46 scalar-normalization/covariance gate. The perturbative term $100(\alpha/\pi)^2$ is the first unevaluated order scale after the declared one-loop witness. Since all residuals lie below the combined non-fit envelope, the residual is assigned to the declared APF/QED covariance channel rather than to a new fitted generation operator.
\end{proof}

\paragraph{No-smuggling audit.}
The following inputs remain forbidden:
\[
\{m_e,m_\mu,m_\tau\}_{\rm target}\to T_\ell,
\quad
r_e,r_\mu,r_\tau,
\quad
\alpha_{\rm chosen\ to\ reduce\ residuals},
\quad
\mu_{\rm chosen\ to\ reduce\ residuals},
\quad
L_{\rm chosen\ to\ reduce\ residuals},
\]
and likewise target-selected thresholds, target-selected matching conventions, and coefficients inferred from the residual vector.

\paragraph{Claim ladder after v48.}
The charged-lepton branch now has
\[
\boxed{
\ell:
[P_{\rm pole\ completion}^{\rm APF\ envelope,CODATA2022}]
+[P_{\rm LO\ QED\ numeric\ transport\ witness}]
+[P_{\rm QED\ truncation\ covariance\ admitted}]
+[P_{\rm full\ QED\ running\ ledger\ typed}/\mathcal L_{\rm QED}^{L\ge2}\ {\rm required}]
}.
\]
Still not claimed:
\[
\boxed{
[P_{\rm QED\ running\ final}],\quad
[P_{\rm physical\ final}],\quad
\Delta_\ell=0,
\quad \text{CODATA-sigma agreement for all leptons}.
}
\]

\paragraph{Next theorem.}
The next theorem is
\[
\boxed{\texttt{APF\_CHARGED\_LEPTON\_FULL\_QED\_NUMERIC\_RUNNING\_WITH\_DECLARED\_EXTERNAL\_COEFFICIENT\_LEDGER}}.
\]
