
% APF v47 charged-lepton CODATA ledger import and leading-order QED transport
% Generated package layer: v47

\section{Charged-Lepton CODATA Ledger Import and Leading-Order QED Transport}
\label{sec:charged-lepton-codata-ledger-lo-qed-v47}

\paragraph{Purpose.}
v47 imports an external constants ledger for the charged-lepton QED boundary and pushes numerical transport as far as is justified without converting QED into a residual-fitting device.  The ledger used here is the CODATA 2022/NIST constants ledger.  The result is a two-part theorem: first, the v46 pole completion is re-audited against the CODATA 2022 charged-lepton references; second, a leading-order, explicitly declared, frozen-\(\alpha\) QED pole-to-running witness is evaluated.  The witness is not a final multi-loop QED running theorem.

\begin{theorem}[CODATA-ledger re-audit of the charged-lepton pole route]
Let the APF v43 pole-codomain vector be
\[
\Phi_\ell^{\rm v43}
=(0.511002635788531,\ 105.658243985342,\ 1776.916832008411)\ {\rm MeV}.
\]
Import the external CODATA 2022 reference ledger
\[
(m_e,m_\mu,m_\tau)_{{\rm CODATA}\,2022}
=(0.51099895069,\ 105.6583755,\ 1776.86)\ {\rm MeV},
\]
with standard uncertainties
\[
(1.6\times 10^{-10},\ 2.3\times 10^{-6},\ 0.12)\ {\rm MeV}.
\]
Then the diagnostic residuals are
\[
(0.000721155792\%,\ -0.000124471588\%,\ 0.003198451674\%).
\]
The APF envelope from v46 is unchanged:
\[
\varepsilon_{\ell,{\rm QED}}^{\rm APF}={1\over 5063},
\qquad
100/5063=0.019751135690302\%.
\]
Hence
\[
\max_i |\delta_i|=0.003198451674\%<0.019751135690302\%,
\qquad
{\max_i |\delta_i|\over 100/5063}=0.161937608245<1.
\]
Therefore the charged-lepton pole completion remains envelope-admitted under the CODATA 2022 ledger:
\[
\boxed{
\ell:[P_{\rm pole\ completion}^{\rm APF\ envelope}]
\quad {\rm survives\ CODATA\ 2022\ ledger\ import.}
}
\]
\end{theorem}

\begin{proof}[Proof sketch]
The CODATA values are imported only after the APF pole vector has already been constructed from the APF trace source, generation operator, count coefficients, scalar normalization, and covariance envelope.  The external values are diagnostic targets, not inputs to the construction.  Direct substitution gives the residual vector above.  Each residual lies inside the predeclared APF envelope \(100/5063\%\), so the v46 pole-completion claim survives the external-ledger update.  This is not a statement that the electron and muon residuals are within CODATA metrological standard uncertainties; it is a statement that the APF pole route is complete inside its own declared QED/scheme covariance envelope.
\end{proof}

\begin{theorem}[Leading-order declared QED running witness]
Declare the leading-order frozen-\(\alpha\) QED conversion witness
\[
\bar m_i^{\rm LO}(\mu_i=m_i)
=
{\Phi_{\ell,i}^{\rm v43}\over 1+\alpha/\pi},
\]
with CODATA 2022
\[
\alpha^{-1}=137.035999177,
\qquad
\alpha=0.007297352564331.
\]
Then
\[
{1\over 1+\alpha/\pi}=0.997682563522362,
\]
and the numerical leading-order QED running witness is
\[
\boxed{
\bar m_\ell^{\rm LO}
=(0.509818419640,\ 105.413387716567,\ 1772.798940124186)\ {\rm MeV}.
}
\]
This is an admissible numerical transport witness because the scheme, scale convention, loop order, and \(\alpha\) input are fixed before evaluation.
\end{theorem}

\begin{corollary}[Boundary of the v47 numerical claim]
v47 does not claim a final QED-running mass vector.  A final running export still requires the full external ledger
\[
\mathcal L_{\rm QED}^{\rm final}
=\{S_{\rm run},\mu,L,\alpha(\mu),\text{threshold schedule},\text{matching convention},\text{mass anomalous dimension},\Sigma_{\rm QED}\}
\]
at the intended perturbative order and convention.  The v47 witness fixes a reproducible first numerical route object; it does not exhaust higher-loop running, vacuum-polarization thresholds, scheme conversion, or covariance propagation.
\end{corollary}

\paragraph{No-smuggling audit.}
The following choices are not allowed: selecting \(\alpha\), \(\mu\), loop order, threshold schedule, matching convention, or scale convention by minimizing the residuals; replacing the common QED operator by three generation-specific factors; or using CODATA masses to reconstruct \(T_\ell\), \(\mathcal R_\ell\), or the APF count coefficients.  The only imported numeric constants are ledger entries, and their role is declared before evaluation.

\paragraph{Status after v47.}
\[
\boxed{
\ell:
[P_{\rm pole\ completion}^{\rm APF\ envelope}\ {\rm under\ CODATA\ 2022}]
+
[P_{\rm LO\ QED\ numeric\ transport\ witness}]
}
\]
Still not claimed:
\[
\boxed{
[P_{\rm QED\ running\ final}],\quad
[P_{\rm physical\ final}],\quad
\Delta_\ell=0.
}
\]
