
% APF v70 -- Charged-Lepton QED Running Evaluator
% Generated 2026-05-10

\section{Charged-Lepton QED Running Evaluator}
\label{sec:v70-charged-lepton-qed-running}

\subsection{Purpose}
The trace-to-scheme export sprint left the charged-lepton sector at
\[
\ell:
[P_{\rm pole\,completion}]^{\rm APF\,envelope,CODATA2022}
+
[P_{\rm QED\,boundary}].
\]
This module closes the source-clean one-loop QED evaluator and keeps the full multi-loop QED running claim fail-closed until a complete external coefficient, threshold, scheme, and covariance ledger is supplied.

\subsection{Frozen APF input}
The charged-lepton pole-codomain APF vector is
\[
\Phi_\ell^{v43}=
(0.511002635789,\;105.658243985342,\;1776.916832008411)\;{\rm MeV}.
\]
The CODATA 2022 pole ledger used for audit is
\[
(m_e,m_\mu,m_\tau)_{\rm CODATA2022}
=
(0.51099895069,\;105.6583755,\;1776.86)\;{\rm MeV}.
\]
The fine structure constant ledger is frozen as
\[
\alpha^{-1}=137.035999177(21),
\qquad
\frac{\alpha}{\pi}=0.002322819464195.
\]

\subsection{Closed evaluator}
The source-clean one-loop same-self-scale evaluator is defined by
\[
\overline m_i^{\rm QED,LO}(\mu_i=m_i)
=
\frac{M_i^{\rm pole}}{1+\alpha/\pi}.
\]
Therefore
\[
\boxed{
\overline m_\ell^{\rm QED,LO}(\mu_i=m_i)
=
(0.509818419641,\;105.413387716567,\;1772.798940124185)\;{\rm MeV}.
}
\]
This is a declared QED transport object, not a new APF trace derivation.

\subsection{Audit residuals remain pole-codomain diagnostics}
Before QED running, the APF pole vector has CODATA residuals
\[
\delta_\ell=
(+0.000721155884\%,\;-0.000124471588\%,\;+0.003198451674\%).
\]
These residuals were already inside the APF/CODATA envelope in the closeout registry. The running evaluator is not used to fit, erase, or redistribute those residuals.

\subsection{No-smuggling rule}
No scale is selected from the residual vector. No coefficient is selected from the residual vector. No separate generation factors
\((r_e,r_\mu,r_\tau)\) are admitted. In particular,
\[
\overline m_i\mapsto r_i\overline m_i,
\qquad
r_i\;{\rm chosen\;from}\;(m_e,m_\mu,m_\tau)_{\rm target},
\]
is forbidden. QED running is a codomain transport operation, not a residual-fitting operation.

\subsection{Fail-closed full-QED boundary}
The full multi-loop running object remains
\[
U_{\rm QED}^{L\ge 2}(
\beta_i,\gamma_m^i,\mathcal T,\mathcal M,\Sigma_{\rm QED}),
\]
where \(\mathcal T\) denotes thresholds, \(\mathcal M\) matching conventions, and \(\Sigma_{\rm QED}\) covariance. Without that external ledger the evaluator must fail closed.

\subsection{Status theorem}
\begin{theorem}[Charged-lepton QED evaluator closeout]
Given the APF pole-complete charged-lepton vector \(\Phi_\ell^{v43}\) and the frozen CODATA 2022 fine-structure constant ledger, the one-loop same-self-scale QED evaluator is closed:
\[
\ell:
[P_{\rm pole\,completion}]^{\rm APF\,envelope,CODATA2022}
+
[P_{\rm QED\,LO\,self\text{-}scale\,evaluator}].
\]
The full multi-loop QED running route remains an external-ledger boundary:
\[
\ell:
[P_{\rm full\,QED\,boundary}]/
[\beta_i,\gamma_m^i,\mathcal T,\mathcal M,\Sigma_{\rm QED}\;{\rm required}].
\]
\end{theorem}

\paragraph{Nonclaims.}
This module does not claim \([P_{\rm physical\,final}]\), exact zero residuals, or a multi-loop QED final mass vector. It closes the source-clean LO evaluator and the fail-closed boundary for all higher-order evaluator claims.
