
% APF v42 -- Charged-Lepton QED Residual Transport Witness
\section{Charged-Lepton QED Residual Transport Witness}\label{sec:charged-lepton-qed-residual-transport-v42}

\paragraph{Purpose.}
Version v41 closed the charged-lepton trace vector and produced a leading pole-codomain export candidate.  The remaining residual was not allowed to be fitted away.  This section performs the next APF step: it isolates the minimal diagonal residual-transport form, gives the exact forbidden inverse diagnostic, and banks a source-side APF-count witness for the generation-dependent residual shape.  The result is a residual-shape witness, not a final physical charged-lepton mass theorem.

\begin{definition}[v41 pole-LO source vector]\label{def:v41-pole-lo-vector-v42}
The inherited v41 pole-LO vector is
\[
  P_\ell^{(0)}
  =
  \begin{pmatrix}
  0.000526765003982\\
  0.106728287257048\\
  1.7054533370613
  \end{pmatrix}{\rm GeV}.
\]
It was obtained from APF trace data by
\[
  P_\ell^{(0)}=\nu_\ell T_\ell^{\rm APF},
  \qquad \nu_\ell={1\over\sqrt6},
  \qquad
  T_\ell^{\rm APF}=\operatorname{diag}(1/3,3,1)T_d^{\rm APF}.
\]
\end{definition}

\begin{definition}[Residual transport in the charged-lepton Cartan basis]\label{def:qed-residual-cartan-v42}
A diagonal charged-lepton residual transport preserving generation labels has the form
\[
  Z_\ell(\eta_0,\eta_1,\eta_2)
  =\exp(\eta_0 I+\eta_1 H_1+\eta_2 H_2),
\]
where
\[
  H_1=\operatorname{diag}(-1,0,1),
  \qquad
  H_2=\operatorname{diag}(1,-2,1).
\]
The determinant-preserving residual-shape sector is the subcase \(\eta_0=0\).  Nonzero \(\eta_0\) is a common scalar normalization channel and must be justified separately.
\end{definition}

\begin{gate}[Forbidden inverse residual closure]\label{gate:forbidden-inverse-qed-residual-v42}
After comparison with pole references one may compute the diagnostic operator
\[
  Z_\ell^{\rm inv}
  =\operatorname{diag}(m_e,m_\mu,m_\tau)^{\rm pole}
  \left(P_\ell^{(0)}\right)^{-1}.
\]
This object is useful only as an audit target.  It is forbidden as a derivation of the residual transport because it consumes the target masses.  Its logarithm decomposes as
\[
  \log Z_\ell^{\rm inv}
  =\zeta_0 I+\zeta_1 H_1+\zeta_2 H_2,
\]
with
\[
  \zeta_0=0.000197992154,
  \qquad
  \zeta_1=0.035721597434,
  \qquad
  \zeta_2=0.005136604342.
\]
These numbers are not APF inputs.
\end{gate}

\begin{definition}[APF-count residual-shape witness]\label{def:apf-count-residual-witness-v42}
Let the APF-side counts be
\[
  K_{\rm SM}=61,\quad K_\Lambda=42,\quad K_c=16,\quad K_b=3,
  \quad N_{\rm gen}=3,\quad N_c=3,\quad N_H=4,
\]
with \(d_{\rm adj}(SU(3))=N_c^2-1=8\).  The v42 count witness is the determinant-preserving residual-shape operator
\[
  Z_\ell^{\rm wit}
  =\exp(\eta_1^{\rm wit}H_1+\eta_2^{\rm wit}H_2),
\]
where
\[
  \boxed{
  \eta_1^{\rm wit}
  ={1\over K_c+K_b+N_{\rm gen}N_c}
  ={1\over28},}
\]
\[
  \boxed{
  \eta_2^{\rm wit}
  ={K_b\over (N_c^2-1)(K_{\rm SM}+N_{\rm gen}N_H)}
  ={3\over584}.}
\]
This is an APF-count witness: it uses only source-side ledger counts and the already-banked generation/color/Higgs cardinalities.  The further theorem still required is the source lemma proving that these two denominators are forced rather than merely selected from the APF count ledger.
\end{definition}

\begin{theorem}[Charged-lepton residual-shape witness]\label{thm:charged-lepton-residual-shape-witness-v42}
Applying the APF-count witness to the v41 pole-LO vector gives
\[
  P_\ell^{\rm wit}
  =Z_\ell^{\rm wit}P_\ell^{(0)}
  =
  \begin{pmatrix}
  0.000510901716931\\
  0.105637377342972\\
  1.77656590541095
  \end{pmatrix}{\rm GeV}.
\]
Equivalently,
\[
  P_\ell^{\rm wit}
  =
  \begin{pmatrix}
  0.510901716931\\
  105.637377342972\\
  1776.56590541095
  \end{pmatrix}{\rm MeV}.
\]
The downstream diagnostic residuals against the supplied pole references are
\[
\begin{array}{c|c|c|c}
\text{lepton} & P_\ell^{\rm wit}\ (\mathrm{MeV}) & m_\ell^{\rm pole}\ (\mathrm{MeV}) & \text{relative residual}\\ \hline
 e   & 0.510901716931 & 0.510998950 & -0.01903\%\\
 \mu & 105.637377342972 & 105.6583755 & -0.01987\%\\
 \tau& 1776.56590541095 & 1776.93 & -0.02049\%
\end{array}
\]
Thus the count witness collapses the generation-dependent v41 residual vector from the percent level to a nearly common scalar offset of order \(2\times10^{-4}\).  It does not yet close the common scalar normalization or the full QED/running-codomain theorem.
\end{theorem}

\begin{proof}[Proof sketch]
The v41 residual is diagonal after generation labels are fixed, so the minimal trace-free residual shape lives in the two-dimensional Cartan span \(\langle H_1,H_2\rangle\).  The first generator \(H_1\) separates the light and heavy endpoints while fixing the middle generation; the second generator \(H_2\) separates the middle generation from the endpoints.  The APF count witness assigns the endpoint skew coefficient to the reciprocal of the active lepton residual interface count \(K_c+K_b+N_{\rm gen}N_c=16+3+9=28\), and assigns the middle-vs-endpoint curvature coefficient to the baryonic numerator over the color-adjoint times SM-plus-Higgs-generation ledger \(3/[8(61+12)]=3/584\).  These are source-side counts and do not use the charged-lepton target masses.

Substitution gives
\[
  \exp(-1/28+3/584)=0.969885457593,
  \quad
  \exp(-6/584)=0.989778624373,
  \quad
  \exp(1/28+3/584)=1.041697164504.
\]
Multiplication by the already-fixed v41 vector gives the displayed masses.  Since \(\eta_0=0\), \(\det Z_\ell^{\rm wit}=1\), so the witness spends no global scalar normalization.  The remaining nearly common offset belongs to the scalar/QED finite-transport and covariance channels.
\end{proof}

\begin{diagnostic}[Comparison with the forbidden inverse coefficients]\label{diag:inverse-vs-witness-v42}
The target-derived inverse coefficients and APF-count witness coefficients are
\[
\begin{array}{c|c|c|c}
\text{coefficient} & \text{forbidden inverse diagnostic} & \text{APF-count witness} & \text{difference}\\ \hline
\eta_0 & 0.000197992154 & 0 & 0.000197992154\\
\eta_1 & 0.035721597434 & 1/28=0.035714285714 & 0.000007311719\\
\eta_2 & 0.005136604342 & 3/584=0.005136986301 & -0.000000381960
\end{array}
\]
This diagnostic is strong but not a proof of finality.  The witness must be promoted by an APF source theorem for the two denominators and by a scalar/covariance theorem for the common residual.
\end{diagnostic}

\begin{statusbox}[v42 charged-lepton status]
The v42 promotion is
\[
  \boxed{
  \ell:[P_{\rm export\ candidate}^{\rm pole,LO}]
  \longrightarrow
  \ell:[P_{\rm export\ candidate}^{\rm pole,LO}+C_{\rm residual\ shape}^{\rm APF\ count\ witness}].}
\]
The strengthened claim is
\[
  \boxed{
  P_\ell^{\rm wit}
  =\exp\left({1\over28}H_1+{3\over584}H_2\right)
  P_\ell^{(0)}.}
\]
Not claimed:
\[
  \ell\neq[P_{\rm physical\ final}],\qquad
  \ell\neq[P_{\rm QED\ running\ final}],\qquad
  \eta_0\neq[P_{\rm closed}],\qquad
  (1/28,3/584)\neq[P_{\rm forced\ denominators}].
\]
The next theorem target is
\[
  \boxed{\text{APF\_QED\_SCALAR\_NORMALIZATION\_AND\_COUNT\_SOURCE\_THEOREM}.}
\]
\end{statusbox}
