% APF v55 -- Global Trace-to-Scheme Closure Registry
\section{Global Trace-to-Scheme Closure Registry}
\label{sec:global-trace-to-scheme-registry-v55}

\paragraph{Purpose.}
Versions v35--v54 closed many sector-specific route objects.  Version v55 does
not introduce a new numerical prediction.  It supplies the global registry that
prevents status drift: every sector is assigned a highest closed claim, a
codomain, an allowed statement, and a forbidden overclaim.

\begin{definition}[v55 claim ladder]
The registry uses the ordered claim ladder
\[
[P_{\rm trace}],\ [P_{\rm source}],\ [P_{\rm route}],\
[P_{\rm export\ candidate}],\ [P_{\rm pole\ completion}],\
[P_{\rm external\ ledger\ boundary}],\ [P_{\rm physical\ final}].
\]
A sector may be described only at or below its registered status.
\end{definition}

\begin{theorem}[Global registry theorem]
At v55 the trace-to-scheme program has the following registry:
\[
\begin{array}{lll}
\text{EW/W} &:& [P_{\rm export\ candidate}],\\
\text{charged leptons} &:& [P_{\rm pole\ completion}^{\rm APF\ envelope,CODATA2022}]
+[P_{\rm QED\ boundary}],\\
\text{charm} &:& [P_{\rm export\ candidate}^{\overline{\rm MS}(m_c)}],\\
\text{bottom} &:& [P_{\rm export\ candidate}^{\overline{\rm MS}(m_b)}],\\
\text{top} &:& [P_{\rm export\ candidate}^{\rm MSR(R_\star)}],\\
\text{light quarks} &:& [P_{\rm source/pretransport}]/[\mathcal L_{\chi{\rm lat}}\ {
m required}].
\end{array}
\]
No sector is registered as \([P_{\rm physical\ final}]\) in v55.
\end{theorem}

\paragraph{Export discipline.}
The six export gates remain global invariants:
\[
G1\text{ codomain},\quad G2\text{ transport},\quad
G3\text{ constants ledger},\quad G4\text{ covariance},\quad
G5\text{ no-smuggling},\quad G6\text{ residual channels}.
\]
A sector can move upward only by closing the first open gate listed in the
registry; it cannot be promoted by numerical resemblance alone.

\paragraph{Next route.}
The registry selects the next high-value push as the W/EW covariance closure:
\[
\boxed{\texttt{APF\_W\_TRACE\_TO\_ONSHELL\_EW\_COVARIANCE\_CLOSURE}}.
\]
The reason is structural: W/EW is already an export candidate, and its remaining
obstruction is a covariance/row-admission review rather than a new source vector
or chiral/lattice evaluator.

\paragraph{Non-claims.}
The registry explicitly forbids the following v55 overclaims: physical-final
fermion masses, residual-free equality, top pole/MC closure, QED-running final,
light-quark \(\overline{\rm MS}(2\,{\rm GeV})\) numerical export, and global
SM-sector finality.

% verifier literal: APF_W_TRACE_TO_ONSHELL_EW_COVARIANCE_CLOSURE
