% APF v53 -- Light-Quark Up Trace Source and Scalar Knockout
\section{Light-Quark Up Trace Source and Scalar-Transport Knockout}
\label{sec:light-quark-source-scalar-v53}

\paragraph{Purpose.}
Version v52 typed the light-quark chiral/lattice transport gate but left the
first up-family light trace source open.  Version v53 closes the source side by
importing the upstream APF trace-sector source theorem for the up family, then
checks whether any identity or universal-scalar route can honestly export the
light traces to the external
\(\overline{\rm MS}(2\,{\rm GeV})\) codomain.  The answer is no.

\begin{theorem}[Light-quark source closure]
The APF trace-sector source bank supplies the full light trace vector
\[
T_{uds}^{\rm APF}
= (T_u,T_d,T_s)
= (0.0011532098523996768,\
   0.003870916422334003,\
   0.08714328163365137)\ {\rm GeV}.
\]
Equivalently,
\[
T_{uds}^{\rm APF}
= (1.153209852399677,\;3.870916422334003,\;87.14328163365137)\ {\rm MeV}.
\]
These entries are source traces.  They are not physical
\(\overline{\rm MS}(2\,{\rm GeV})\) masses.
\end{theorem}

\begin{theorem}[Identity and universal-scalar knockout]
Let \(m_{uds}^{\overline{\rm MS}}(2\,{\rm GeV})\) denote the external light-quark
codomain.  Neither the identity map
\[
T_{uds}^{\rm APF}\mapsto T_{uds}^{\rm APF}
\]
nor any single universal scalar map
\[
T_{uds}^{\rm APF}\mapsto \kappa T_{uds}^{\rm APF}
\]
constitutes an APF-admissible export map to this codomain.  Identity misses the
central external ledger values by
\[
\delta_u=-46.610654981496\%,\qquad
\delta_d=-17.640076120553\%,\qquad
\delta_s=-6.798629268822\%.
\]
The best unweighted least-squares scalar is
\[
\kappa_{\rm LS}=1.073363492801811,
\]
but it still leaves
\[
(-42.693826152538\%,\;-11.597864437866\%,\;+0.038948821934\%).
\]
Thus the missing object is not a scalar normalization but a genuine low-energy
chiral/lattice transport evaluator.
\end{theorem}

\paragraph{Ratio diagnostic.}
The APF source ratios are
\[
T_u/T_d=0.297916494850162,
\qquad
T_s/T_d=22.5123128804490,
\qquad
T_s/\bigl((T_u+T_d)/2\bigr)=34.6899249216306.
\]
These are trace-source ratios, not already the physical chiral ratios of the
\(\overline{\rm MS}(2\,{\rm GeV})\) codomain.

\paragraph{No-smuggling rule.}
The up trace may not be defined by inverting \(m_u(2\,{\rm GeV})\).  The
chiral/lattice transport map may not be chosen by least-squares fit to
\((m_u,m_d,m_s)\).  The lattice scale, chiral ansatz, EM/isospin convention,
renormalization factor \(Z_m\), and covariance matrix must be declared before
comparison.

\paragraph{Status.}
Version v53 changes the light-quark branch from a missing-source obstruction to
a source-closed transport obstruction:
\[
\boxed{
(u,d,s):[P_{\rm source}^{T_{uds}}]
+[P_{\rm identity/scalar\ knockout}]
/[\mathcal L_{\chi{\rm lat}}+\Sigma_{uds}\ {\rm required}].
}
\]
It does not claim a light-quark export candidate or final physical masses.
