% APF v54 -- Light-Quark APF-Side Completion Boundary
\section{Light-Quark APF-Side Completion Boundary}
\label{sec:light-quark-apf-boundary-v54}

\paragraph{Purpose.}
After v53 the light-quark source vector is no longer missing:
\[
T_{uds}^{\rm APF}=(1.153209852399677,3.870916422334003,87.14328163365137)\ {\rm MeV}.
\]
The remaining question is whether APF alone can export this vector to
\((m_u,m_d,m_s)^{\overline{\rm MS}}(2\,{\rm GeV})\).  Version v54 closes the
answer: APF has completed the source and pretransport side, but the physical
light-quark codomain is defined through an external chiral/lattice-QCD
evaluator with its own constants, calibration choices, and covariance.

\begin{theorem}[No APF-only light-quark export]
Given the source-closed vector \(T_{uds}^{\rm APF}\), an APF-only map to
\(\overline{\rm MS}(2\,{\rm GeV})\) is inadmissible unless it supplies a
source-independent chiral/lattice evaluator
\[
U_{\chi{\rm lat}}(\mathcal L_{\chi{\rm lat}},\Sigma_{uds}).
\]
Identity transport and universal scalar transport are already knocked out, and
a diagonal component-wise map solved from the three light-quark targets is
inverse fitting.  Therefore the APF-side branch is complete at the boundary
\[
\boxed{
T_{uds}^{\rm APF}
\quad/\quad
\mathcal L_{\chi{\rm lat}}+\Sigma_{uds}\ {\rm required}.
}
\]
\end{theorem}

\paragraph{Required external evaluator ledger.}
The next admissible object is not another APF scalar.  It is an external route
ledger containing hadronic calibration data, lattice spacing and action,
chiral extrapolation/interpolation prescription, EM/isospin convention,
renormalization factor \(Z_m\), conversion to
\(\overline{\rm MS}(2\,{\rm GeV})\), and a joint covariance matrix.

\paragraph{Status.}
Version v54 closes the APF-side light-quark branch as
\[
\boxed{
(u,d,s):[P_{\rm source}^{T_{uds}}]
+[P_{\rm pretransport\ complete}]
/[\mathcal L_{\chi{\rm lat}}+\Sigma_{uds}\ {\rm external}].
}
\]
It does not claim a numerical light-quark export candidate.  The next valid
push must either import a declared source-independent chiral/lattice evaluator
or stop at this boundary.
