% APF v52 -- Light-Quark Chiral/Lattice Transport Gate
\section{Light-Quark Chiral--Lattice Transport Gate}
\label{sec:light-quark-chiral-lattice-v52}

\paragraph{Purpose.}
The heavy-quark routes close because their codomains are either self-scale
$\overline{\rm MS}$ masses or an explicitly typed MSR scale.  The light quarks
are different.  The external reference quantities
\[
(m_u,m_d,m_s)^{\overline{\rm MS}}(2\,{\rm GeV})
\]
are low-energy current-quark masses extracted through chiral and lattice-QCD
machinery.  A trace anchor is therefore not exportable by identity transport.

\begin{theorem}[Light-quark transport gate]
Let
\[
T_{uds}^{\rm APF}=(T_u,T_d,T_s)
\]
be the APF light-quark trace source vector.  A physical light-quark export to
$\overline{\rm MS}(2\,{\rm GeV})$ is APF-admissible only if there exists a
source-independent chiral/lattice transport evaluator
\[
U_{\chi{\rm lat}}(\mathcal L_{\chi{\rm lat}}):
T_{uds}^{\rm APF}\longmapsto
(m_u,m_d,m_s)^{\overline{\rm MS}}(2\,{\rm GeV}),
\]
where
\[
\mathcal L_{\chi{\rm lat}}=
\{a,\text{scale setting},\text{lattice action},Z_m,
\text{chiral ansatz},\text{EM/isospin convention},
\alpha_s,\text{thresholds},\Sigma_{uds}\}.
\]
The evaluator and its constants ledger may not be solved from the external
$(m_u,m_d,m_s)$ targets.
\end{theorem}

\paragraph{Source inventory.}
The inherited down-family trace source contains
\[
T_d^{\rm APF}=0.003870916422334\ {\rm GeV},\qquad
T_s^{\rm APF}=0.087143281633652\ {\rm GeV},
\]
but v52 does not contain an APF-native first-generation up trace source
$T_u^{\rm APF}$.  The heavy up-family anchors $T_c$ and $T_t$ do not determine
$T_u$ without an additional APF source theorem.

\paragraph{Identity-transport knockout.}
The identity map is explicitly rejected.  Using the down-family trace components
as if they were already $\overline{\rm MS}(2\,{\rm GeV})$ values gives
\[
T_d^{\rm APF}=3.870916422334\ {\rm MeV},\qquad
T_s^{\rm APF}=87.143281633652\ {\rm MeV},
\]
whereas the external light-quark codomain ledger uses approximately
\[
m_d(2\,{\rm GeV})=4.70\ {\rm MeV},\qquad
m_s(2\,{\rm GeV})=93.5\ {\rm MeV}.
\]
The diagnostic residuals are
\[
\delta_d=-17.640076120553\%,\qquad
\delta_s=-6.798629269891\%.
\]
They are not repaired by rescaling the trace anchors; they are assigned to the
missing chiral/lattice transport channel.

\paragraph{No-smuggling rule.}
The following are forbidden: defining $T_u$ by inverting $m_u(2\,{\rm GeV})$,
choosing a diagonal three-component correction from the PDG values, selecting a
lattice scale or chiral ansatz by best APF agreement, or using heavy-quark
self-scale routes as substitutes for low-energy chiral transport.

\paragraph{Status.}
Version v52 closes the codomain declaration, the typed transport schema, the
source inventory, the identity-transport knockout, and the no-smuggling audit:
\[
\boxed{
(u,d,s):[P_{\rm light\ transport\ gate}^{\chi{\rm lat}}]
/[T_u+\mathcal L_{\chi{\rm lat}}+\Sigma_{uds}\ {\rm required}].
}
\]
It does not claim a light-quark export candidate or final physical masses.
