
\section{v76: QCD Coefficient Table Vendoring and Fixed-Flavour Four-Loop Evaluator}

This module upgrades the charm--bottom QCD evaluator from the v73 LO scaffold to a source-vendored fixed-flavour four-loop running evaluator.  It is deliberately narrower than full RunDec parity: no threshold crossing is performed in the closed v76 route.  The closed routes are
\[
 m_c^{\overline{\rm MS}}(m_c)\to m_c^{\overline{\rm MS}}(3\,{\rm GeV}),
 \qquad
 m_b^{\overline{\rm MS}}(m_b)\to m_b^{\overline{\rm MS}}(10\,{\rm GeV}),
\]
with fixed active flavours \(n_f=4\) and \(n_f=5\), respectively.

\paragraph{Vendored source layer.}
The coefficient table \texttt{qcd\_coefficient\_table\_v76.json} vendors the four-loop fixed-flavour \(\alpha_s\) running coefficients and the RunDec-style mass-running functions
\[
 {m(\mu)\over m(\mu_0)}={c_{n_f}(\alpha_s(\mu)/\pi)\over c_{n_f}(\alpha_s(\mu_0)/\pi)}.
\]
The table records the source family as Chetyrkin--Kuehn--Steinhauser RunDec, Herren--Steinhauser RunDec/CRunDec v3, and the four-loop QCD beta-function lineage.

\paragraph{Numerical results.}
Using the APF self-scale inputs
\[
 m_c(m_c)=1.272334177712\, {\rm GeV},\qquad \alpha_s(m_c)=0.38,
\]
\[
 m_b(m_b)=4.177490455927\, {\rm GeV},\qquad \alpha_s(m_b)=0.223,
\]
v76 obtains
\[
 \boxed{m_c^{\rm v76}(3\,{\rm GeV})=0.992321292393\,{\rm GeV}},
 \qquad
 \boxed{\alpha_s^{\rm v76}(3\,{\rm GeV})=0.250671393527},
\]
\[
 \boxed{m_b^{\rm v76}(10\,{\rm GeV})=3.637640144573\,{\rm GeV}},
 \qquad
 \boxed{\alpha_s^{\rm v76}(10\,{\rm GeV})=0.177208097457}.
\]

\paragraph{Claim status.}
\[
 \boxed{c,b:[P_{\rm 4loop\ fixed\text{-}n_f\ evaluator\ closed}]}
\]
while
\[
 \boxed{[P_{\rm full\ RunDec\ threshold\ parity}]\ \text{is still not claimed}.}
\]
The remaining full-RunDec step is threshold/matching validation against a real RunDec/CRunDec adapter or published benchmark examples.  v76 is therefore a real evaluator closure, but not a full physical-final mass claim.
