\section{RunDec Adapter or Coefficient-Vendoring Gate (v74)}
\label{sec:v74-rundec-adapter-vendoring}

\paragraph{Purpose.}
Version v73 closed an executable leading-order QCD running scaffold and a fail-closed adapter interface.  Version v74 closes the next engineering/theory gate: a RunDec-compatible external evaluator may enter the APF trace-to-scheme export stack only through either (i) a validated adapter, or (ii) a vendored coefficient table satisfying a typed manifest.  No silent replacement of the evaluator by hand-entered target masses is permitted.

\paragraph{Objects inherited from v71--v73.}
The APF self-scale objects are
\[
 m_c^{\overline{\rm MS}(m_c),{\rm APF}}=1.272334177712\ {\rm GeV},
 \qquad
 m_b^{\overline{\rm MS}(m_b),{\rm APF}}=4.177490455927\ {\rm GeV}.
\]
The v73 source-clean leading-order scaffold gives
\[
 m_c^{\rm LO}(3\,{\rm GeV})=1.070790477803\ {\rm GeV},
 \qquad
 \alpha_s^{\rm LO}(3\,{\rm GeV})=0.265307067708,
\]
\[
 m_b^{\rm LO}(10\,{\rm GeV})=3.737911682909\ {\rm GeV},
 \qquad
 \alpha_s^{\rm LO}(10\,{\rm GeV})=0.180200450083.
\]
These numbers remain a baseline self-test, not a full RunDec result.

\paragraph{The v74 gate.}
A full QCD evaluator is admissible only if one of the following certificates is supplied.

\begin{enumerate}
\item \textbf{Validated adapter certificate.}  An external command declared by \texttt{RUNDEC\_CMD} must accept the APF JSON payload, return charm and bottom results, declare loop order, active-flavor schedule, thresholds, matching convention, covariance policy, and validation identifier, and state that no target mass was consumed as an input.
\item \textbf{Vendored coefficient certificate.}  A coefficient file declared by \texttt{RUNDEC\_COEFFICIENT\_JSON} must supply the required coefficient families
\[
 \beta_i,
 \qquad
 \gamma_m^i,
 \qquad
 \zeta_g^i,
 \qquad
 \zeta_m^i,
 \qquad
 d_i^{\rm OS/\overline{MS}},
\]
plus source identifiers, loop order, normalization convention, active-flavor convention, threshold convention, and validation examples.
\end{enumerate}

\paragraph{No-smuggling theorem.}
The adapter or coefficient table is rejected if it includes any of the comparison targets
\[
 m_c^{\rm PDG},\quad m_b^{\rm PDG},\quad
 m_c(3\,{\rm GeV})_{\rm target},\quad m_b(10\,{\rm GeV})_{\rm target}
\]
as calibration inputs.  They may appear only in an audit block, after transport has been evaluated.

\paragraph{Fail-closed result.}
The v74 package includes a runnable gate script.  In the present package no external RunDec command and no vendored coefficient table are present.  Therefore the full evaluator correctly returns
\[
 \boxed{\texttt{FULL\_RUNDEC\_EVALUATOR\_FAIL\_CLOSED}.}
\]
This is a positive result: it proves the stack will not silently use an incomplete evaluator or tune QCD running to the observed charm and bottom masses.

\paragraph{Status.}
Version v74 promotes the charm/bottom external-evaluator layer to
\[
 \boxed{
 c,b:
 [P_{\rm adapter/vendor\ gate}^{\rm RunDec}]
 /
 [P_{\rm full\ numeric\ evaluator}\ {\rm pending\ validated\ adapter\ or\ coefficients}].
 }
\]
It does not claim a full multi-loop QCD result.

\paragraph{Next module.}
The next step is no longer conceptual.  It is either
\[
 \boxed{\texttt{APF\_v75\_RUNDEC\_ADAPTER\_VALIDATION}}
\]
if an executable RunDec/CRunDec command is supplied, or
\[
 \boxed{\texttt{APF\_v75\_QCD\_COEFFICIENT\_TABLE\_VENDORING}}
\]
if the coefficient tables are vendored directly.
