
% APF v72 -- RunDec Source Ledger and Full QCD Evaluator Gate
\section{RunDec Source Ledger and Full QCD Evaluator Gate}
\label{sec:rundec-source-ledger-full-qcd-gate-v72}

\paragraph{Purpose.}
The preceding charm--bottom QCD evaluator closed only the leading-order running map.  This module freezes the authoritative external source ledger required to replace that leading-order evaluator by a full QCD running, decoupling, matching, and mass-conversion evaluator.  The point is not to rederive perturbative QCD inside APF.  It is to make the external evaluator boundary exact, source-controlled, and fail-closed.

\begin{definition}[RunDec/CRunDec external evaluator]
The external QCD evaluator for charm and bottom self-scale $\overline{\rm MS}$ transport is the RunDec/CRunDec family of implementations, with v3 as the active source ledger.  The evaluator is admissible for APF export only when the code path declares the active flavour number, threshold convention, loop order, decoupling prescription, mass anomalous dimension, beta-function order, and covariance protocol.
\end{definition}

\begin{definition}[Full-QCD evaluator route]
For $q\in\{c,b\}$, the full external route has the typed form
\[
  \Phi_q^{\overline{\rm MS}}(\mu_1)
  = U_{\rm RunDec}\!igl(\mathcal L_{\rm RunDec},\Sigma_q\bigr)
    \left[m_q^{\overline{\rm MS}}(m_q)_{\rm APF}\right],
\]
where
\[
  \mathcal L_{\rm RunDec}
  =\{\alpha_s, n_f, \beta_i, \gamma_m^i, \zeta_g^i,\zeta_m^i,
      \mu_0,\mu_1,\mathcal T, L, \mathcal P_{\rm match},\mathcal C\}.
\]
Here $\mathcal T$ is the threshold schedule, $L$ is the loop order, $\mathcal P_{\rm match}$ is the matching prescription, and $\mathcal C$ is the source manifest for all coefficient tables.
\end{definition}

\begin{theorem}[RunDec source-ledger closure]
\label{thm:rundec-source-ledger-closure-v72}
The full QCD evaluator source ledger for charm and bottom is closed when the following source families are named and typed:
\begin{enumerate}
\item RunDec/CRunDec v3 for running and decoupling of $\alpha_s$ and quark masses, including five-loop beta-function and four-loop decoupling support;
\item the original RunDec implementation paper for the canonical Mathematica lineage and formula collection;
\item CRunDec for the C++ implementation lineage;
\item the five-loop QCD beta-function source;
\item the four-loop on-shell/$\overline{\rm MS}$ mass-relation source for heavy-quark mass scheme transformations.
\end{enumerate}
These sources close the source ledger for a full external evaluator, but they do not by themselves certify a local APF implementation unless the coefficient tables or an audited RunDec-compatible engine are actually ingested and tested.
\end{theorem}

\begin{proof}[Proof sketch]
RunDec/CRunDec is precisely the external tool family built for the required operation: running and decoupling of the strong coupling and quark masses, together with heavy-quark mass conversion formulae.  The v3 paper records the higher-order upgrades needed beyond the leading-order APF scaffold.  The original RunDec and CRunDec papers provide implementation lineage.  The beta-function and mass-relation sources identify the coefficient families needed for higher-order evaluation.  Therefore the external source ledger is complete.  However, APF's no-smuggling rule forbids treating the existence of a source as equivalent to a computed value; a local evaluator must either call/import a validated implementation or fail closed.
\end{proof}

\begin{theorem}[Full-QCD evaluator gate]
\label{thm:full-qcd-evaluator-gate-v72}
The charm--bottom route may be promoted from
\[
  [P_{\rm LO\ QCD\ running\ evaluator}]
\]
to
\[
  [P_{\rm full\ RunDec\ evaluator\ closed}]
\]
only if a local evaluator satisfies all of the following gates:
\begin{enumerate}
\item all coefficient families required by the declared loop order are ingested from the source ledger;
\item the active-flavour and threshold schedule is declared;
\item decoupling and matching conventions are declared;
\item uncertainty and covariance are propagated from mass, $\alpha_s$, threshold, truncation, and implementation channels;
\item the evaluator is checked against published RunDec/CRunDec examples or an equivalent validation set;
\item target masses at the destination scale are not used to choose scales, thresholds, or coefficients.
\end{enumerate}
If any gate fails, the route remains
\[
  [P_{\rm RunDec\ source\ ledger\ closed}/P_{\rm implementation\ pending}].
\]
\end{theorem}

\paragraph{v72 closeout.}
This module closes the source ledger and evaluator gate, not the full numeric evaluator.  The correct v72 status is
\[
  c,b:
  [P_{\rm RunDec\ source\ ledger\ closed}]
  +[P_{\rm full\ QCD\ evaluator\ gate}]
  /[P_{\rm coefficient\ ingestion\ pending}].
\]
The v71 leading-order evaluator remains valid as a scaffold, but it is not a substitute for a RunDec-compatible higher-order computation.

\paragraph{No-smuggling audit.}
The following moves are forbidden: selecting $\mu_1$ from the desired output mass, changing thresholds to reduce residuals, omitting decoupling uncertainties, mixing 90\% confidence intervals with one-sigma covariance without declaring the conversion, and calling a RunDec reference a local numerical evaluation before the actual code path is ingested or validated.

\paragraph{Next module.}
The next engineering module is
\[
  \texttt{APF\_v73\_RUNDEC\_COMPATIBLE\_QCD\_EVALUATOR\_IMPLEMENTATION}.
\]
It should either vendor/call a RunDec-compatible implementation or ingest the coefficient tables explicitly and validate against published examples.
