\section{RunDec Adapter Validation Harness (v75)}
\label{sec:v75-rundec-adapter-validation}

\paragraph{Purpose.}
Versions v72--v74 isolated the full charm/bottom QCD evaluator boundary: a complete RunDec/CRunDec-compatible evaluator is admissible only through a validated external adapter or a vendored coefficient table.  Version v75 closes the adapter-validation harness itself.  It does not vendor RunDec and it does not claim full multi-loop running.  It proves that any future adapter must pass a typed protocol before its numbers can enter the APF export registry.

\paragraph{Inherited APF objects.}
The charm and bottom APF self-scale objects remain
\[
 m_c^{\overline{\rm MS}(m_c),{\rm APF}}=1.272334177712\ {\rm GeV},
 \qquad
 m_b^{\overline{\rm MS}(m_b),{\rm APF}}=4.177490455927\ {\rm GeV}.
\]
The v73 internal scaffold remains only a leading-order baseline,
\[
 m_c^{\rm LO}(3\,{\rm GeV})=1.070790477803\ {\rm GeV},
 \qquad
 m_b^{\rm LO}(10\,{\rm GeV})=3.737911682909\ {\rm GeV}.
\]
These values are not promoted to a full RunDec result.

\paragraph{Adapter protocol.}
A RunDec-compatible command must accept a JSON request with the charm and bottom source objects, declared start and target scales, active-flavor settings, and forbidden-input list.  It must return a JSON response containing
\[
\{\text{results},\text{loop order},\text{threshold schedule},\text{matching convention},\text{covariance policy},\text{validation id},\text{no-target-reuse flag}\}.
\]
The validator rejects responses that omit any field, include target masses as inputs, fail to identify loop order or thresholds, or fail to declare source-independent evaluation.

\paragraph{Protocol self-test.}
The package contains a deliberately synthetic protocol adapter used only to test the validator.  The synthetic adapter is marked
\[
 \boxed{\texttt{protocol\_self\_test\_only=true}}
\]
and is rejected for physics promotion.  This prevents the validation harness from being mistaken for a RunDec implementation.

\paragraph{Environment audit.}
The package checks for a real executable adapter through \texttt{RUNDEC\_CMD}.  In the present environment no validated RunDec command is supplied.  Therefore the physical full-evaluator route correctly remains
\[
 \boxed{\texttt{FULL\_RUNDEC\_EVALUATOR\_FAIL\_CLOSED}.}
\]

\paragraph{Status.}
Version v75 promotes the external QCD layer to
\[
 \boxed{
 c,b:
 [P_{\rm RunDec\ adapter\ validation\ harness\ closed}]
 /
 [P_{\rm full\ numeric\ evaluator}\ {\rm pending\ real\ adapter\ or\ vendored\ coefficients}].
 }
\]
It does not claim
\[
 [P_{\rm full\ RunDec\ evaluator}],\qquad
 [P_{\rm full\ multi\text{-}loop\ QCD\ final}],\qquad
 [P_{\rm physical\ final}].
\]

\paragraph{Next module.}
The remaining work is now concrete implementation work:
\[
 \boxed{\texttt{APF\_v76\_QCD\_COEFFICIENT\_TABLE\_VENDORING}}
\]
or, if a working external binary/library is supplied,
\[
 \boxed{\texttt{APF\_v76\_RUNDEC\_ADAPTER\_BINARY\_VALIDATION}}.
\]
