
\section{APF v61: Top MSR Coefficient Ingestion and Four-Loop Audit}
\label{sec:top-msr-v61}

\paragraph{Purpose.}
Version v60 implemented a fail-closed evaluator for the APF top object
\[
  m_t^{\rm MSR}(R_\star)=168.1690557938\,\mathrm{GeV},\qquad
  R_\star=85.857222698385\,\mathrm{GeV}.
\]
Version v61 ingests sourced coefficients for the MSR-to-pole series through four-loop order, evaluates the resulting pole witness, and audits the result before any claim promotion.

\paragraph{Coefficient normalization.}
The evaluator uses
\[
  m_t^{\rm pole}-m_t^{\rm MSR}(R)
  = R\sum_{n=1}^L c_n\left(\frac{\alpha_s(R)}{4\pi}\right)^n.
\]
The ingested coefficient vector is
\[
  (c_1,c_2,c_3,c_4)
  = \left(
  5.333333333333,\ 131.785000000000,\ 4699.703000000000,\ 214829.905730030296
  \right),
\]
where the fourth coefficient is obtained from the top four-loop on-shell/MSbar coefficient
\((8.615\pm0.017)\alpha_s^4\) after conversion to the \(\alpha_s/(4\pi)\) normalization.

\paragraph{Four-loop running witness.}
Using \(n_f=5\) QCD beta coefficients through \(\beta_3\), the evaluator gives
\[
  \alpha_s(R_\star)=0.119084611158869.
\]
The termwise mass shifts are
\[
\begin{aligned}
  \Delta_1 &= 4.339316649273\,\mathrm{GeV},\\
  \Delta_2 &= 1.016095141009\,\mathrm{GeV},\\
  \Delta_3 &= 0.343387582223\,\mathrm{GeV},\\
  \Delta_4 &= 0.148749217342\,\mathrm{GeV}.
\end{aligned}
\]
Thus
\[
  \Phi_t^{\rm pole,4\ell}
  = 174.016604383647\,\mathrm{GeV}.
\]

\paragraph{Audit result.}
Against the audit-only direct/MC-style top reference
\(172.56\pm0.31\,\mathrm{GeV}\), the residual is
\[
  \Delta_t^{4\ell}=1.456604383647\,\mathrm{GeV}.
\]
A conservative envelope including the quoted audit uncertainty, \(\alpha_s\) parametric uncertainty, the four-loop coefficient uncertainty, pole-renormalon floor, and five-loop-and-beyond estimate is
\[
  \sigma_{v61}=0.449435570715\,\mathrm{GeV},
  \qquad
  \frac{\Delta_t^{4\ell}}{\sigma_{v61}}=3.240963729972.
\]
Therefore the sourced four-loop pole witness does not support promotion to physical-final top closure.

\begin{center}
\fbox{\parbox{0.92\linewidth}{
\textbf{v61 conclusion.} The top branch is now stronger but more honest: coefficient ingestion and a four-loop executable evaluator are closed, while top physical-final closure is blocked. The earlier LO near-hit is not stable under sourced higher-order QCD coefficients. Top remains an APF MSR-route object unless a revised APF-native codomain or full R-evolution comparison theorem is supplied.
}}
\end{center}

\paragraph{No-smuggling rule.}
The PDG/direct top value is audit-only. It is not used to select \(R_\star\), \(\alpha_s\), loop order, or any coefficient. Coefficients are fixed before comparison. If agreement worsens at higher order, the result is reported rather than tuned away.
