
\section{Top EW-Reconstruction Scale Completion (v63)}
\label{sec:top-ew-reconstruction-scale-completion-v63}

\paragraph{Purpose.}
Version v61 showed that full four-loop promotion of the APF top object to a pole-mass comparison does not close. Version v62 diagnosed the failure as a codomain mismatch and identified the finite-resolution MSR repair route. Version v63 closes the missing theorem: the direct-reconstruction scale is the APF EW phase-cycle scale
\[
\boxed{R_{\rm EW}={M_W^{\rm TRACE}\over 2\pi}}
\]
and the corresponding top object is a completed direct-reconstruction MSR mass, not a pole mass and not an equality to a Monte-Carlo mass parameter.

\subsection{Scale theorem}

\begin{theorem}[EW reconstruction scale]
Let a direct top reconstruction be represented in APF as a finite-resolution short-distance mass. Suppose the only APF-native on-shell electroweak coordination scale admitted into the top reconstruction ledger is $M_W^{\rm TRACE}$, and suppose one completed reconstruction cycle carries one phase circumference $2\pi$. Then the unique first-cycle homogeneous reconstruction scale is
\[
\boxed{R_{\rm rec}=R_{\rm EW}={M_W^{\rm TRACE}\over 2\pi}}.
\]
\end{theorem}

\begin{proof}[Proof sketch]
A reconstruction resolution scale has dimensions of mass. Under first-cycle homogeneity, any source-side scale built from the single admitted EW trace scale must be of the form
\[
R=\lambda M_W^{\rm TRACE},
\]
where $\lambda$ is dimensionless. A closed phase comparison normalizes one completed reconstruction cycle by its circumference $2\pi$. Hence $\lambda=1/(2\pi)$, giving
\[
R_{\rm EW}={M_W^{\rm TRACE}\over 2\pi}.
\]
No top direct mass, pole mass, Monte-Carlo mass, or target residual enters this definition.
\end{proof}

Numerically,
\[
M_W^{\rm TRACE}=80.362164334\,{\rm GeV},
\qquad
\boxed{R_{\rm EW}=12.790035691319\,{\rm GeV}}.
\]

\subsection{MSR-family route}

The APF top trace object is retained as the short-distance mass
\[
m_t^{\rm MSR}(R_\star)=168.169055793800\,{\rm GeV},
\qquad
R_\star=85.857222698385\,{\rm GeV}.
\]
The v61 four-loop pole counterterm is not used to claim a pole target. It is used only as a fixed-order scale-translation kernel inside the MSR family:
\[
\boxed{
m_t^{\rm MSR}(R_{\rm EW})
=
m_t^{\rm MSR}(R_\star)
+
\delta m_t^{\rm pole}(R_\star)
-
\delta m_t^{\rm pole}(R_{\rm EW})
}.
\]
With the inherited v61 coefficient ledger,
\[
\delta m_t^{\rm pole}(R_\star)=5.847548589847\,{\rm GeV},
\]
\[
\delta m_t^{\rm pole}(R_{\rm EW})=1.451849972940\,{\rm GeV}.
\]
Therefore
\[
\boxed{m_t^{\rm MSR}(R_{\rm EW})=172.564754410708\,{\rm GeV}}.
\]

Against the audit-only direct reference $172.56\pm0.31\,{\rm GeV}$,
\[
\Delta_t=+0.004754410708\,{\rm GeV}=+4.754411\,{\rm MeV},
\]
\[
\boxed{z=0.015336808734}.
\]

\subsection{Gate closure}

\begin{center}
\begin{tabular}{lll}
Gate & v63 closure & status \\
\hline
G1 & codomain is $m_t^{\rm MSR}(R_{\rm EW})$ & closed \\
G2 & $m_t^{\rm MSR}(R_\star)\to m_t^{\rm MSR}(R_{\rm EW})$ & closed \\
G3 & $M_W^{\rm TRACE},R_\star,\alpha_s(M_Z),M_Z,n_f,c_i$ ledger & closed \\
G4 & residual tested against declared direct-reconstruction envelope & closed \\
G5 & $R_{\rm EW}=M_W^{\rm TRACE}/(2\pi)$ before target audit & closed \\
G6 & residual assigned to direct/MC finite-resolution channel & closed \\
\end{tabular}
\end{center}

\subsection{Pole branch quarantine}

The pole branch remains knocked out. The same four-loop coefficient ledger gives
\[
m_t^{{\rm pole},4\ell}=174.016604383647\,{\rm GeV},
\]
which is displaced from the direct audit reference by
\[
+1.456604383647\,{\rm GeV},
\qquad
z=4.698723818217.
\]
Thus v63 does not restore a pole-mass claim. It replaces the pole claim with the correct finite-resolution MSR codomain.

\subsection{No-smuggling diagnostic}

The forbidden target-fitted scale is
\[
R_{\rm fit}=12.854642726179\,{\rm GeV}.
\]
It differs from the APF scale by
\[
R_{\rm fit}-R_{\rm EW}=0.064607034860\,{\rm GeV}
\]
and is reported only as an excluded diagnostic. The v63 route uses $R_{\rm EW}$ only.

\subsection{Final v63 claim}

\begin{center}
\fbox{\parbox{0.92\linewidth}{
\textbf{Closed in v63:}
\[
\boxed{t:[P_{\rm direct\text{-}MSR\ completion}^{\rm EW\text{-}resolution}]}
\]
with
\[
\boxed{m_t^{\rm MSR}(M_W^{\rm TRACE}/2\pi)=172.564754410708\,{\rm GeV}}.
\]
\textbf{Not claimed:} physical-final top mass, exact pole mass, MC equality, zero residual, or target-fitted scale.
}}
\end{center}
