
\section{Top Codomain Diagnosis and MSR-Resolution Repair (v62)}
\label{sec:top-codomain-diagnosis-msr-repair-v62}

\paragraph{Purpose.}
Version v61 closed an important negative result: the APF top object does not survive promotion to a full four-loop pole-mass comparison against the direct/MC-style top-mass audit.  The purpose of v62 is to diagnose whether this is a failure of the APF trace anchor or a failure of codomain selection.

\begin{center}
\fbox{\parbox{0.92\linewidth}{
\textbf{v62 conclusion.}  The v61 failure is best understood as a codomain mismatch.  The direct top mass is a kinematic/Monte-Carlo-style mass parameter, not a clean pole-mass observable.  The pole branch remains knocked out, but an MSR finite-resolution branch is opened and evaluated.
}}
\end{center}

\subsection{The v61 pole-branch knockout}

The v61 evaluator used the sourced four-loop MSR--pole counterterm
\[
\delta m_t^{\rm pole}(R)
=R\sum_{n=1}^{4}c_n\left({\alpha_s(R)\over 4\pi}\right)^n
\]
with the same coefficient ledger as v61.  At the APF scale
\[
R_\star=85.857222698385\ {\rm GeV},
\qquad
m_t^{\rm MSR}(R_\star)=168.169055793800\ {\rm GeV},
\]
the four-loop pole witness is
\[
\delta m_t^{\rm pole}(R_\star)=5.847548589847\ {\rm GeV},
\]
so that
\[
 m_t^{\rm pole,4\ell}=174.016604383647\ {\rm GeV}.
\]
Against the audit-only direct top reference
\[
 m_t^{\rm dir}=172.56\pm0.31\ {\rm GeV},
\]
this gives
\[
\Delta_t^{\rm pole}=+1.456604383647\ {\rm GeV}.
\]
Thus the direct/MC pole-promotion branch remains knocked out.

\subsection{Codomain repair: direct reconstruction as a finite-resolution MSR mass}

The repair is not to tune the pole conversion.  The repair is to compare the APF top object to the correct codomain.  Direct reconstruction measures a finite-resolution top-mass parameter.  The corresponding APF route should therefore remain in the MSR family and translate from the APF scale $R_\star$ to a reconstruction scale $R_{\rm rec}$:
\[
 m_t^{\rm MSR}(R_\star)
 \longrightarrow
 m_t^{\rm MSR}(R_{\rm rec}).
\]
At fixed order the sourced pole counterterm may be used as a scale-translation kernel without claiming the pole mass as the target:
\[
\boxed{
 m_t^{\rm MSR}(R_{\rm rec})
 =
 m_t^{\rm MSR}(R_\star)
 +\delta m_t^{\rm pole}(R_\star)
 -\delta m_t^{\rm pole}(R_{\rm rec}).
 }
\]
This cancels the pole object from the comparison and keeps the target in a short-distance finite-resolution mass scheme.

\subsection{EW-resolution witness}

The APF-native EW reconstruction scale is taken to be
\[
\boxed{
 R_{\rm EW}={M_W^{\rm TRACE}\over 2\pi}
 }
\]
with
\[
 M_W^{\rm TRACE}=80.362164334\ {\rm GeV},
\qquad
 R_{\rm EW}=12.790035691319\ {\rm GeV}.
\]
Using the v61 four-loop coefficient ledger,
\[
\alpha_s(R_{\rm EW})=0.168529007262,
\]
\[
\delta m_t^{\rm pole}(R_{\rm EW})=1.451849972940\ {\rm GeV}.
\]
Therefore
\[
\boxed{
 m_t^{\rm MSR}(R_{\rm EW})
 =172.564754410708\ {\rm GeV}.
 }
\]
The residual against the audit-only direct reference is
\[
\Delta_t^{\rm EW-MSR}=+0.004754410708\ {\rm GeV}
=4.754410708\ {\rm MeV},
\]
with
\[
\Delta_t^{\rm EW-MSR}/0.31=0.015336808734.
\]

\subsection{No-smuggling audit}

The target-fitted scale solving
\[
 m_t^{\rm MSR}(R_{\rm fit})=172.56\ {\rm GeV}
\]
is
\[
R_{\rm fit}=12.854642726179\ {\rm GeV}.
\]
This quantity is reported only as a forbidden diagnostic.  It is not used to define the route.  The route scale
\[
 R_{\rm EW}=M_W^{\rm TRACE}/(2\pi)
\]
is selected from the already banked EW trace sector and does not consume the direct top-mass target.

\subsection{Claim status}

\begin{center}
\fbox{\parbox{0.92\linewidth}{
\textbf{Closed:} the v61 pole failure is diagnosed as a codomain mismatch for direct/MC auditing; the pole branch is quarantined.\\[2mm]
\textbf{Conditional witness:} the finite-resolution MSR route with $R_{\rm EW}=M_W/(2\pi)$ gives $m_t^{\rm MSR}(R_{\rm EW})=172.564754410708$ GeV.\\[2mm]
\textbf{Not claimed:} physical-final top mass, exact pole mass, MC equality, target-fitted scale, or full R-evolution finality.
}}
\end{center}

The next theorem must either prove the EW-resolution scale $R_{\rm EW}$ as the correct direct-reconstruction scale or replace it with an externally calibrated MC/MSR scale ledger.
