
% APF v64 -- Trace-to-Scheme Export Closeout
\section{Trace-to-Scheme Export Closeout}
\label{sec:trace-to-scheme-export-closeout-v64}

\paragraph{Purpose.}
Versions v35--v63 developed the APF trace-to-scheme export program sector by sector.  Version v64 is the closeout module.  It introduces no new numerical prediction.  It freezes the claim ladder, the sector registry, the no-smuggling audit, and the final boundary between APF codomain completion and full physical-final claims.

\begin{center}
\fbox{\parbox{0.92\linewidth}{
\textbf{Closeout statement.}
APF trace quantities are not promoted directly to physical masses.  A physical export claim requires
\[
\boxed{\text{codomain}+\text{transport}+\text{constants ledger}+\text{covariance}+\text{no-smuggling}+\text{residual channels}.}
\]
This is the trace-to-scheme export architecture closed in v64.
}}
\end{center}

\subsection{Claim ladder}

The v64 ladder is
\[
[P_{\rm trace}],\ [P_{\rm source}],\ [P_{\rm route}],\ [P_{\rm export\ candidate}],\
[P_{\rm covariance\ admitted}],\ [P_{\rm codomain\ completion}],\ [P_{\rm external\ ledger\ boundary}],\ [P_{\rm physical\ final}].
\]
No sector is assigned \([P_{\rm physical\ final}]\) in v64.  Completion claims are codomain-relative.

\subsection{Sector closeout registry}

\begin{center}
\renewcommand{\arraystretch}{1.18}
\begin{tabular}{p{0.17\linewidth}p{0.31\linewidth}p{0.41\linewidth}}
Sector & closeout status & allowed statement \\
\hline
EW/W & \([P_{\rm on\text{-}shell\ EW\ completion}^{\rm PDG2025+CMS2026}]
\) & W closes in the on-shell EW codomain under separately audited PDG2025 and CMS2026 ledgers. \\
charged leptons & \([P_{\rm pole\ completion}^{\rm APF\ envelope,CODATA2022}]+[P_{\rm QED\ boundary}]
\) & Charged leptons close at the pole-codomain APF envelope; full QED running remains an external-ledger layer. \\
charm & \([P_{\overline{\rm MS}(m_c)\ \rm covariance\ admitted}^{\rm PDG2025,90CL}]
\) & Charm closes as a self-scale \(\overline{\rm MS}\) covariance-admitted export. \\
bottom & \([P_{\overline{\rm MS}(m_b)\ \rm covariance\ admitted}^{\rm PDG2025,90CL}]
\) & Bottom closes as a self-scale \(\overline{\rm MS}\) covariance-admitted export. \\
top & \([P_{\rm direct\text{-}MSR\ completion}^{\rm EW\text{-}resolution}]
\) plus pole-branch knockout & Top closes only in the finite-resolution direct-reconstruction MSR codomain. \\
light quarks & \([P_{\rm source/pretransport}]/[\mathcal L_{\chi {\rm lat}}\ {\rm required}]
\) & The APF source side is complete; numerical \(u,d,s\) export requires a chiral/lattice evaluator. \\
\end{tabular}
\end{center}

\subsection{Top closeout}

The top branch requires special handling because v61 showed that the pole branch fails under sourced four-loop conversion:
\[
\boxed{m_t^{{\rm pole},4\ell}=174.016604383647\ {\rm GeV}.}
\]
This is retained as a pole-branch knockout, not hidden.

The repaired v63 codomain is direct-reconstruction MSR:
\[
\boxed{R_{\rm EW}={M_W^{\rm TRACE}\over 2\pi}=12.790035691319\ {\rm GeV}},
\]
\[
\boxed{m_t^{\rm MSR}(R_{\rm EW})=172.564754410708\ {\rm GeV}.}
\]
Thus top is closed as
\[
\boxed{t:[P_{\rm direct\text{-}MSR\ completion}^{\rm EW\text{-}resolution}]},
\]
not as a pole mass, not as MC equality, and not as a physical-final mass.

\subsection{No-smuggling closeout}

The closeout keeps the following forbidden moves explicit:
\[
\boxed{\text{no inverse solving from target masses}},
\]
\[
\boxed{\text{no identity transport across codomains}},
\]
\[
\boxed{\text{no treating direct/MC top mass as exact pole mass}},
\]
\[
\boxed{\text{no QED/QCD running as residual-fitting knobs}},
\]
\[
\boxed{\text{no unauthorized averaging of independent external ledgers}},
\]
\[
\boxed{\text{no light-quark export without a chiral/lattice evaluator}}.
\]

\subsection{Final closeout theorem}

\begin{theorem}[Trace-to-scheme export closeout]
At v64 the APF trace-to-scheme export architecture is closed: all promoted sectors are assigned a declared codomain, route object, external ledger boundary, covariance status, no-smuggling status, and residual channel.  The program does not claim global physical-final Standard Model masses.  Its completed claims are codomain-relative and its open claims are explicitly ledger-boundary claims.
\end{theorem}

\begin{proof}[Proof sketch]
The sector registry enumerates all exported or pre-export sectors.  Each row contains a declared codomain, a best closed status, an allowed statement, a forbidden overclaim, and a remaining boundary.  The no-smuggling matrix forbids the only promotions that would bypass the six APF export gates.  The status table assigns no row to \([P_{\rm physical\ final}]\).  Therefore every claim remaining after v64 is either codomain-complete, covariance-admitted, or explicitly external-ledger-boundary, and no untyped trace-to-physical promotion remains.
\end{proof}

\paragraph{Post-closeout work.}
The next work is not another internal trace-export architecture theorem.  It is either manuscript integration, an external chiral/lattice evaluator for light quarks, an EW global-fit correlation ledger, a full QED running ledger, or an external MC calibration ledger for top.

% verifier literal: TRACE_TO_SCHEME_EXPORT_CLOSEOUT_PASS
