% APF v56 -- W Trace to On-Shell EW Covariance Closure
\section{W Trace to On-Shell Electroweak Covariance Closure}
\label{sec:w-onshell-ew-covariance-closure-v56}

\paragraph{Purpose.}
Version v56 closes the next registry-selected route object: the W trace anchor
is promoted from an export candidate to an on-shell electroweak completion under
separately declared external covariance ledgers.  The closure is deliberately
narrow.  It concerns the on-shell W mass branch, not the entire electroweak
sector and not an unauthorized global average.

\begin{definition}[v56 on-shell W audit ledgers]
The APF trace anchor is
\[
M_W^{\rm APF\text{-}TRACE}=80.362164334000\,{\rm GeV}.
\]
The external audit ledgers are frozen separately:
\[
M_W^{\rm PDG2025}=80.3692\pm 0.0133\,{\rm GeV},
\qquad
M_W^{\rm CMS2026}=80.3602\pm 0.0099\,{\rm GeV}.
\]
The PDG and CMS ledgers are comparison ledgers only; neither is used to construct
the APF trace anchor, the route map, or the residual envelope.
\end{definition}

\begin{theorem}[W on-shell covariance completion]
The W trace branch satisfies all six export gates at the on-shell covariance
audit level.  Its residuals are
\[
\Delta_{\rm PDG2025}=-0.007035666000\,{\rm GeV}
= -7.035666\,{\rm MeV},
\qquad
z_{\rm PDG2025}=-0.528997443610,
\]
\[
\Delta_{\rm CMS2026}=0.001964334000\,{\rm GeV}
= 1.964334\,{\rm MeV},
\qquad
z_{\rm CMS2026}=0.198417575757.
\]
Hence
\[
\max(|z_{\rm PDG2025}|,|z_{\rm CMS2026}|)=0.528997443610<1.
\]
Therefore the W route is promoted to
\[
\boxed{W:[P_{\rm on\text{-}shell\ EW\ completion}^{\rm PDG2025+CMS2026\ ledgers}]}.
\]
\end{theorem}

\begin{proof}[Audit proof]
Gate G1 is closed because the codomain is the on-shell W mass.  Gate G2 is the
same-input on-shell EW route inherited from v35.  Gate G3 is closed at this
layer by separately declaring the PDG 2025 and CMS 2026 comparison ledgers.  Gate
G4 is closed because the APF residual is evaluated in each ledger's stated
uncertainty model and lies within one standard deviation in both ledgers.  Gate
G5 is closed because the external masses are not used in the APF construction.
Gate G6 is closed because the nonzero residuals are assigned to the declared
external covariance/measurement ledgers and inherited on-shell EW transport
channel, rather than absorbed into the trace anchor.  Thus the route is complete
at the declared on-shell covariance level.
\end{proof}

\paragraph{Non-combination rule.}
Version v56 does not combine the PDG 2025 and CMS 2026 ledgers.  Such a
combination requires an external correlation and working-group protocol.  The
audit is therefore a two-ledger robustness check, not a new world average.

\paragraph{Non-claims.}
The theorem does not claim zero residual, all electroweak observables, the full
EW global fit, or physical-final Standard Model closure.  It closes only the
on-shell W mass branch under the declared ledgers.
