
\section{Light-Quark Trace Obstruction Map}\label{sec:light-quark-obstruction-map}

This section closes the light-quark branch at the trace/obstruction level.  The APF trace anchors are
\[
 m_u^{\rm APF\mbox{-}TRACE}=1.1532098524~{\rm MeV},\qquad
 m_d^{\rm APF\mbox{-}TRACE}=3.8709164223~{\rm MeV},\qquad
 m_s^{\rm APF\mbox{-}TRACE}=87.1432816337~{\rm MeV}.
\]
The comparison target is not a pole or self-scale mass.  The relevant external lane is the conventional current-quark
\(\overline{\rm MS}\) normalization at \(\mu=2~{\rm GeV}\):
\[
 m_u(2~{\rm GeV})=2.16\pm0.07~{\rm MeV},\qquad
 m_d(2~{\rm GeV})=4.70\pm0.07~{\rm MeV},\qquad
 m_s(2~{\rm GeV})=93.5\pm0.8~{\rm MeV},
\]
with uncertainties quoted at 90\% confidence level in the PDG summary.

\begin{center}
\begin{tabular}{lrrrr}
\hline
quark & APF TRACE [MeV] & PDG \(\overline{\rm MS}(2\,{\rm GeV})\) [MeV] & residual [MeV] & quoted-scale pull \\
\hline
\(u\) & 1.1532098524 & 2.16 & -1.0067901476 & -14.38 \\
\(d\) & 3.8709164223 & 4.70 & -0.8290835777 & -11.84 \\
\(s\) & 87.1432816337 & 93.5 & -6.3567183663 & -7.95 \\
\hline
\end{tabular}
\end{center}

Thus direct export to the light-quark \(\overline{\rm MS}(2~{\rm GeV})\) codomain is knocked out.  The ratio lane does not rescue the export:
\[
 (m_u/m_d)_{\rm APF}=0.2979165,\qquad (m_s/m_d)_{\rm APF}=22.5123,
\qquad (m_s/\hat m)_{\rm APF}=34.6899.
\]
These are not jointly consistent with the PDG light-quark ratio lane.

\begin{center}
\fbox{\begin{minipage}{0.86\linewidth}
\textbf{Claim boundary.}
The light-quark branch is closed as
\[
(u,d,s)^{\rm APF\mbox{-}TRACE}:[P_{\rm trace}+P_{\rm low\mbox{-}energy\ QCD\ obstruction}],
\]
not as a physical export.  The first failed gate is
\[
\texttt{APF\_LIGHT\_QUARK\_TRACE\_TO\_MSBAR\_2GEV\_CHIRAL\_LATTICE\_TRANSPORT\_MAP}.
\]
\end{minipage}}
\end{center}

The obstruction is physical rather than cosmetic: light-quark masses are current-quark parameters inferred through low-energy QCD, chiral symmetry, lattice matching, and conventional scale normalization.  APF trace anchors may still be meaningful internal-sector values, but a physical light-quark export requires an explicit chiral/lattice/QCD transport theorem.
