\section{Electroweak Trace Sector Closure and On-Shell Export Boundary}
\label{sec:ew-sector-closure-v22}

This section closes the electroweak trace sector at the level appropriate for the present paper.  The closure claim is deliberately narrower than a first-principles reconstruction of the full Standard Model electroweak loop stack.  The closed object is the APF trace-sector package: the weak-angle trace anchor, its coupling-ratio consequence, the W trace anchor, a reviewed same-input on-shell transport validation, and an explicit scheme-boundary table.

\subsection{Weak-angle trace anchor}

APF fixes the electroweak trace anchor
\begin{equation}
  \sin^2\theta_W^{\rm APF\text{-}TRACE}=\frac{3}{13}=0.230769230769\ldots .
\end{equation}
Consequently
\begin{equation}
  \cos^2\theta_W^{\rm APF\text{-}TRACE}=\frac{10}{13},
  \qquad
  \tan^2\theta_W^{\rm APF\text{-}TRACE}=\frac{3}{10}.
\end{equation}
At the trace level this yields the coupling-ratio consequence
\begin{equation}
  \boxed{\frac{g'}{g}=\sqrt{\frac{3}{10}}=0.547722557505\ldots .}
\end{equation}
Equivalently,
\begin{equation}
  \frac{g^2}{g^2+g'^2}=\frac{10}{13},
  \qquad
  \frac{g'^2}{g^2+g'^2}=\frac{3}{13}.
\end{equation}
These are trace-sector coupling-geometry statements, not scheme-free claims about running electroweak couplings.

\subsection{W trace anchor and reviewed same-input transport}

The APF W trace anchor is
\begin{equation}
  M_W^{\rm APF\text{-}TRACE}=80.362164334\ {\rm GeV}.
\end{equation}
The reviewed same-input DIZET transport evaluation at the APF route point gives
\begin{equation}
  M_W^{\rm DIZET}(x_{\rm APF})=80.357341077578084\ {\rm GeV}.
\end{equation}
Thus
\begin{equation}
  M_W^{\rm DIZET}(x_{\rm APF})-M_W^{\rm APF\text{-}TRACE}
  =-4.823256\ {\rm MeV}.
\end{equation}
With the row/input covariance and DIZET theory-toggle scale used in the route ledger,
\begin{equation}
  \sigma_{\rm input+theory}(M_W)=4.424866\ {\rm MeV},
\end{equation}
so the displacement is
\begin{equation}
  \frac{|\Delta M_W|}{\sigma}=1.090034.
\end{equation}
The proper status is therefore
\begin{equation}
  \boxed{M_W^{\rm APF\to OS}:[P_{\rm export\ candidate}],}
\end{equation}
not a claim that APF has independently derived the complete electroweak loop stack.

\subsection{Channel ledger inherited from the W route}

At the APF input point, the admitted DIZET internal channel ledger closes as
\begin{align}
  \Delta r_{\Delta\alpha} &= +0.05907386039640014,\\
  \Delta r_{\rm rem} &= +0.011667933872161376,\\
  \Delta r_{\rho\text{-}{\rm cross}} &= -0.03424000860914665,
\end{align}
with
\begin{equation}
  0.05907386039640014
  +0.011667933872161376
  -0.03424000860914665
  =0.036501785659414865.
\end{equation}
This equals the same-input DIZET total \(\Delta r\).  The ledger is used as a reviewed-code transport ledger, not as an APF-native analytic loop derivation.

\subsection{W/Z bridge and scheme boundary}

Using the external on-shell value
\begin{equation}
  M_Z=91.1876\ {\rm GeV},
\end{equation}
the APF W trace anchor gives the bridge ratio
\begin{equation}
  \frac{M_W^{\rm APF\text{-}TRACE}}{M_Z}=0.881283906299,
\end{equation}
and hence
\begin{equation}
  1-\left(\frac{M_W^{\rm APF\text{-}TRACE}}{M_Z}\right)^2
  =0.223338676499.
\end{equation}
This is an external-\(M_Z\) on-shell bridge, not an APF derivation of \(M_Z\) and not the same object as the trace anchor \(3/13\).  The gap between the trace weak-angle anchor, the on-shell bridge, and the \(\overline{\rm MS}\)/effective weak-angle conventions is part of electroweak scheme transport.

\subsection{Claim ladder}

\begin{center}
\begin{tabular}{lll}
\hline
Object & Status & Meaning \\
\hline
\(\sin^2\theta_W=3/13\) & \([P_{\rm trace}]\) & EW mixing trace anchor \\
\(g'/g=\sqrt{3/10}\) & \([P_{\rm derived}]\) & coupling-ratio consequence \\
\(M_W^{\rm APF\text{-}TRACE}\) & \([P_{\rm trace}]\) & W trace anchor \\
\(M_W^{\rm APF\to OS}\) & \([P_{\rm export\ candidate}]\) & reviewed same-input DIZET route \\
\(M_Z\) & external input & used only for W/Z bridge \\
\(M_W/M_Z\) bridge & \([P_{\rm bridge}]\) & on-shell consistency check \\
Full EW loop derivation & \([OPEN]\) & downstream theorem program \\
\hline
\end{tabular}
\end{center}

The sector-level conclusion is
\begin{equation}
\boxed{\text{The APF electroweak trace sector closes at mixing/coupling-ratio/W-export-candidate level.}}
\end{equation}
Full first-principles Standard Model loop reconstruction remains a downstream APF-native theorem program; it is not a prerequisite for electroweak trace-sector closure.
