
\section{Fermion Trace Sector Closure and Route-Dependent Export Boundary}
\label{sec:fermion-sector-master-closure}

This section consolidates the APF fermion-sector results into a single route-governed closure statement.  The claim is not that every APF trace entry is already a physical mass in a conventional scheme.  The claim is sharper and safer: the fermion trace sector is locally closed, while physical export is controlled by explicit scheme and transport gates.

\begin{center}
\fbox{%
\begin{minipage}{0.92\linewidth}
\textbf{Master claim.}  The fermion trace sector is locally closed.  Physical mass export is route-dependent.  At the present level, the bottom self-scale \(\overline{\rm MS}\) route reaches export-candidate status; the top, charged-lepton, charm, and light-quark branches close as named transport or codomain obstructions.
\end{minipage}}
\end{center}

\subsection{Master outcome table}

\begin{center}
\begin{tabular}{llll}
\hline
Sector & APF object & Status & First boundary \\
\hline
bottom & \(m_b\) & \([P_{\rm export\ candidate}]\) & publication review of \(\overline{\rm MS}\) codomain admission \\
top & \(m_t\) & \([P_{\rm trace}+P_{\rm ext\text{-}MSR}+P_{\rm terminal}] & APF-native MSR scale selection theorem \\
charged leptons & \(e,\mu,\tau\) & \([P_{\rm trace}+P_{\rm norm\ candidate}+P_{\rm residual}] & generation-residual operator derivation \\
charm & \(m_c\) & \([P_{\rm trace}+P_{\overline{\rm MS}\ knockout}] & charm normalization/transport map \\
light quarks & \(u,d,s\) & \([P_{\rm trace}+P_{\rm low\text{-}energy\ QCD}] & chiral/lattice \(\overline{\rm MS}(2\,\mathrm{GeV})\) transport map \\
\hline
\end{tabular}
\end{center}

\subsection{Export-candidate branch}

The only present fermion mass export candidate is the bottom self-scale \(\overline{\rm MS}\) branch:
\[
 m_b^{\rm APF\to\overline{MS}}(m_b)=4.1774904559270665\,\mathrm{GeV}.
\]
Against the standard self-scale \(\overline{\rm MS}\) comparison anchor, the residual is \(-5.509544\,\mathrm{MeV}\), or approximately \(-0.787\sigma\) on the quoted scale.  This is admitted only as an export candidate, not as a final physical derivation.

\subsection{Terminal or obstructed branches}

The top trace value,
\[
 m_t^{\rm APF\text{-}TRACE}=168.1690557938\,\mathrm{GeV},
\]
occupies a high-scale short-distance lane compatible with an external MSR codomain pass, but direct/MC, pole, self-scale \(\overline{\rm MS}\), and electroweak substitution routes are rejected.  A no-smuggling physical export requires an APF-native top trace-to-MSR scale-selection theorem independent of an external pole anchor.

The charged-lepton trace vector is internally closed.  A six-channel normalization candidate \(\lambda_6=1/\sqrt{6}\) reduces the direct pole residuals to the percent level, but a generation-residual operator is not derived.  Therefore the branch closes as a normalization-candidate/residual-obstruction result rather than an export candidate.

The charm trace value,
\[
 m_c^{\rm APF\text{-}TRACE}=0.6800912249261419\,\mathrm{GeV},
\]
is far from the self-scale \(\overline{\rm MS}\) charm target and direct export is rejected.  The remaining object is a charm trace-to-\(\overline{\rm MS}\) normalization or transport map.

For \(u,d,s\), direct \(\overline{\rm MS}(2\,\mathrm{GeV})\), pole, self-scale, constituent, and simple-ratio exports are rejected.  The obstruction is not merely numerical; it is structural, because light-quark current masses are low-energy QCD/chiral/lattice objects.

\subsection{No-smuggling boundary}

The master closure preserves the no-smuggling discipline.  External values are used as comparison anchors or codomain contracts only.  Failed routes are retained as negative controls rather than absorbed into fitted transports.  This prevents a local APF trace closure from being mislabeled as a universal physical mass export.

\begin{center}
\fbox{%
\begin{minipage}{0.92\linewidth}
\textbf{Safe conclusion.}  APF closes the fermion trace sector as a route-governed sector.  Bottom reaches \(\overline{\rm MS}\) self-scale export-candidate status.  Top, charged leptons, charm, and light quarks are not failures of the trace vector; they are resolved into explicit scheme/transport gates.
\end{minipage}}
\end{center}
