
\section{Denner--Sirlin Formula Import and Partial Native One-Loop Assembly}\label{sec:wtrace-v19-denner-import}

The previous coefficient-map layer made the missing object precise: an APF-native one-loop \(\Delta r\) evaluator requires reviewed Denner--Sirlin source formulae, not fitted coefficients.  The present layer converts that obstruction into a source-formula import matrix and evaluates the rows that are already defensible at the APF route point.

\paragraph{Evaluated native rows.}
At the APF same-input point, the admitted running-alpha row is
\[
\Delta r_{\Delta\alpha}=0.05907386039640014,
\]
and the leading top--rho branch is evaluated natively as
\[
\Delta r_{\rho,\mathrm{lead}}
=-\frac{c_W^2}{s_W^2}\frac{3G_Fm_t^2}{8\sqrt{2}\pi^2}
=-0.032452969615.
\]
Thus the partial APF-native one-loop assembly is
\[
\Delta r_{\mathrm{native,partial}}
=0.026620890781.
\]

\paragraph{Target closure.}
The DIZET same-input total is
\[
\Delta r_{\mathrm{DIZET}}(x_{\rm APF})=0.036501785659414865.
\]
The remaining finite/counterterm target is therefore
\[
\Delta r_{\mathrm{ct,target}}=0.009880894878.
\]
Equivalently,
\[
\Delta r_{\Delta\alpha}
+\Delta r_{\rho,\mathrm{lead}}
+\Delta r_{\mathrm{ct,target}}
=\Delta r_{\mathrm{DIZET}}(x_{\rm APF}),
\]
with zero numerical residual at the precision recorded in the verifier.

\paragraph{No-smuggling guard.}
The finite target is not promoted to an APF-native formula.  It is a same-input target functional assembled from the DIZET remainder and rho-resummation/cross correction.  The rows still requiring reviewed formula import are the W and Z transverse self-energies, charge and \(\gamma Z\) field/mixing counterterms, weak-angle counterterm coefficients, the muon-decay vertex and box finite parts, and the tadpole/ghost/gauge-cancellation terms.

\paragraph{Status.}
The W route remains an on-shell export candidate backed by reviewed same-input DIZET transport and now a partial APF-native one-loop assembly.  Full APF-native one-loop closure remains open behind the gate
\[
\boxed{\texttt{APF\_NATIVE\_FULL\_DENNER\_SIRLIN\_FORMULA\_IMPORT\_NOT\_COMPLETE}}.
\]
