
\section{Top-sector final closeout: external MSR codomain without physical export}
\label{sec:top-final-closeout}

The top branch is now closed at the scheme-taxonomy level. The APF trace anchor is
\begin{equation}
  m_t^{\rm APF\mbox{-}TRACE}=168.1690557938\,\mathrm{GeV}.
\end{equation}
It does not admit a direct pole/Monte-Carlo identification and it does not admit a self-scale
\(\overline{\rm MS}\) identification. Relative to the direct/MC-like context
\(172.52\pm0.33\,\mathrm{GeV}\), the residual is
\begin{equation}
  m_t^{\rm APF\mbox{-}TRACE}-m_t^{\rm dir}=-4.350944\,\mathrm{GeV}
  \simeq -13.18\sigma,
\end{equation}
so direct export is knocked out. Relative to the self-scale \(\overline{\rm MS}\) context
\(\overline m_t(\overline m_t)\simeq 162.69\,\mathrm{GeV}\), the trace is higher by
\begin{equation}
  5.479056\,\mathrm{GeV},
\end{equation}
so the ordinary self-scale \(\overline{\rm MS}\) export is also knocked out.

The clean remaining codomain is the high-scale MSR short-distance lane. Using the external one-loop MSR relation
\begin{equation}
  m_t^{\rm MSR}(R)=m_t^{\rm pole}-R\,{4\over3}{\alpha_s(R)\over\pi},
\end{equation}
with the external pole-context input, the APF trace is admitted at
\begin{equation}
  R_*=85.860076078678\,\mathrm{GeV},\qquad
  m_t^{\rm MSR}(R_*)=168.169055793800\,\mathrm{GeV}.
\end{equation}
At the nearby \(R=80\,\mathrm{GeV}\) lane,
\begin{equation}
  m_t^{\rm MSR}(80\,\mathrm{GeV})=168.426519820767\,\mathrm{GeV},
  \qquad
  m_t^{\rm APF\mbox{-}TRACE}-m_t^{\rm MSR}(80\,\mathrm{GeV})=-0.257464026967\,\mathrm{GeV}.
\end{equation}

This closes the top sector as an external-scheme codomain pass, not as a physical mass export. The scale
\(R_*\) is inferred using the external pole context; APF has not yet supplied a native theorem selecting this
MSR scale from trace structure alone. The safe terminal status is
\begin{equation}
  m_t^{\rm APF\mbox{-}TRACE}:
  [P_{\rm trace}+P_{\rm ext\mbox{-}MSR\ codomain}+P_{\rm terminal\ scheme\ closeout}],
\end{equation}
not
\begin{equation}
  m_t^{\rm APF\to physical}:[P_{\rm export}].
\end{equation}
The remaining theorem gate is
\begin{equation}
  \texttt{APF\_NATIVE\_TOP\_TRACE\_TO\_MSR\_SCALE\_SELECTION\_THEOREM\_WITHOUT\_EXTERNAL\_POLE\_ANCHOR}.
\end{equation}
Thus the top branch is finished for the present sector-closure program: direct physical export is rejected,
self-scale \(\overline{\rm MS}\) export is rejected, and the high-scale MSR codomain is admitted only as an
external-scheme pass.
